### 2023-03-27 temp_code/MS_isoviz_code.Rmd
MS junc to iso code

---
title: "Megan's code for Isoviz"
output: html_notebook
---

# 230327
Write a function that maps leafcutter cluster data back to isoforms 
- use psl/gtf, advantage of psl is it is compatible with long read generated gtf which lacks UTR info, it will likely become increasingly common to use long read data to determine which isoforms are present in a specific cell type
- input gene name and cell type, start with RBFOX2 and A375
- output is clusters with counts and linked isoforms
- print image filtered for cluster with <=4 junctions but print pdf with all
- if junctions doesn't map back, have a parameter to include/not include as novel
- go back and only plot isoforms with junction supporting data
- eventually also want to call Andrews predictions and get best predicted guide for junction
- could we even have a screen design setting? to scale it up?

```{r}
# libraries
require(readr)
require(dplyr)
require(tidyr)
require(magrittr)
require(ggplot2)
require(data.table)
require(stringr)
require(ggpubr)
require(grid)
require(gridExtra)
require(cowplot)

```

```{r}
# generate leafcutter cluster files and store for standard cell types
# this is publicly available A375 data used to design the screen

# get leafcutter data in proper format
clusters <- read_tsv("~/isoviz/data-raw/SRR6515912_out-regtools_.junc.a375_m30_p0.01_.sorted", col_names = FALSE, skip = 1) # 15,901
clusters %<>% 
  separate(X1, into = c("coords", "counts"), sep = " ") %>% 
  separate(counts, into = c("junc.counts", "cluster.counts"), sep = "/", convert = TRUE) %>%
  separate(coords, into = c("chr", "start", "end", "name"), sep = ":", convert = TRUE) %>% 
  separate(name, into = c("same", "cluster.n", "extra"), sep = "_") %>% 
  select(-same, -extra)

# remove clusters with no counts, not sure why these are here to begin with- only 7
# filter out chrM and chrY clusters
clusters %<>% filter(cluster.counts > 0, chr != "chrY", chr != "chrM") # 15,886

leafcutter_clusters = as.data.table(clusters)

save(leafcutter_clusters, file = "~/isoviz/data/leafcutter_clusters.rda")
```

```{r}
# could make this a function
# create gene name conversion file from gtf
gtf = read_tsv("~/isoviz/data-raw/gencode.v41.basic.annotation.gtf", col_names = FALSE, skip = 5) %>% filter(X3 == "transcript")
gtf$transcript_id = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[2]])
gtf$transcript_id = sapply(gtf$transcript_id, gsub, pattern = "transcript_id", replacement = "")
gtf$transcript_id = sapply(as.character(gtf$transcript_id), gsub, pattern = '"', replacement = '')
gtf$transcript_id = sapply(as.character(gtf$transcript_id), gsub, pattern = ' ', replacement = '')

gtf$gene_id = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[1]])
gtf$gene_id = sapply(gtf$gene_id, gsub, pattern = "gene_id", replacement = "")
gtf$gene_id = sapply(as.character(gtf$gene_id), gsub, pattern = '"', replacement = '')
gtf$gene_id = sapply(as.character(gtf$gene_id), gsub, pattern = ' ', replacement = '')

gtf$transcript_name = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[6]])
gtf$transcript_name = sapply(gtf$transcript_name, gsub, pattern = "transcript_name", replacement = "")
gtf$transcript_name = sapply(as.character(gtf$transcript_name), gsub, pattern = '"', replacement = '')
gtf$transcript_name = sapply(as.character(gtf$transcript_name), gsub, pattern = ' ', replacement = '')

gtf$gene_name = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[4]])
gtf$gene_name = sapply(gtf$gene_name, gsub, pattern = "gene_name", replacement = "")
gtf$gene_name = sapply(as.character(gtf$gene_name), gsub, pattern = '"', replacement = '')
gtf$gene_name = sapply(as.character(gtf$gene_name), gsub, pattern = ' ', replacement = '')

gtf$gene_type = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[3]])
gtf$gene_type = sapply(gtf$gene_type, gsub, pattern = "gene_type", replacement = "")
gtf$gene_type = sapply(as.character(gtf$gene_type), gsub, pattern = '"', replacement = '')
gtf$gene_type = sapply(as.character(gtf$gene_type), gsub, pattern = ' ', replacement = '')

gtf$transcript_type = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[5]])
gtf$transcript_type = sapply(gtf$transcript_type, gsub, pattern = "transcript_type", replacement = "")
gtf$transcript_type = sapply(as.character(gtf$transcript_type), gsub, pattern = '"', replacement = '')
gtf$transcript_type = sapply(as.character(gtf$transcript_type), gsub, pattern = ' ', replacement = '')

gtf %>% dplyr::select(gene_id, trans_id = transcript_id, gene_name, transcript_name, gene_type, transcript_type) %>% distinct() %>%
  write_tsv("~/isoviz/data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)

```

```{r}
# modify psl file data to include other gene and transcript names
# currently using gencode basic anntotation file
convert = read_tsv("~/isoviz/data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)

load("~/isoviz/data/iso_exon_data.rda") # 839,796
iso_data_complete = iso_exon_data %>% left_join(convert, by = c("gene_id", "trans_id"))

save(iso_data_complete, file = "~/isoviz/data/iso_data_complete.rda")

```

```{r}
# make psl with intron coords instead of exon
# adapted from genomedata.R

mydataset <- fread("~/isoviz/data-raw/gencode.v41.basic.annotation.psl")

mydataset$gene_id = sapply(mydataset$V10, function(x){strsplit(x, "_")[[1]][2]})
mydataset$trans_id = sapply(mydataset$V10, function(x){strsplit(x, "_")[[1]][1]})

# Get coordinates for each "block"
mydataset$blocksizes = sapply(mydataset$V19, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
mydataset$blockstarts = sapply(mydataset$V21, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
mydataset$strand = mydataset$V9
mydataset$chr = mydataset$V14
mydataset$start = mydataset$V16
mydataset$end = mydataset$V17
mydataset$transcript_length = mydataset$end - mydataset$start

# Clean up and seperate blocks into rows
# need to do different adjustments here to get intron coords that will line up with leafcutter
mydataset = mydataset %>% dplyr::select(chr, start, end, trans_id, gene_id,strand, blocksizes, blockstarts,
                            transcript_length) %>% separate_rows(blocksizes, blockstarts)
mydataset$blockstarts=as.numeric(mydataset$blockstarts)
mydataset$blocksizes=as.numeric(mydataset$blocksizes)
mydataset$blockends = mydataset$blockstarts + mydataset$blocksizes

transcript_ids = unique(mydataset$trans_id)
length(transcript_ids) # 116,566
trans_id = c()
intron_starts = c()
intron_ends = c()

for(id in transcript_ids){
  data = mydataset %>% filter(trans_id == id)
  new_ends = data$blockstarts[-1] + 1
  len = nrow(data)
  new_starts = data$blockends[-len]
  n = len-1
  trans_id = c(trans_id, rep(id, n))
  intron_starts = c(intron_starts, new_starts)
  intron_ends = c(intron_ends, new_ends)
}

# Important note: this no longer has single exon transcripts
intron_data = data.frame(trans_id, intron_starts, intron_ends) # 723,230

convert = read_tsv("~/isoviz/data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)
trans_info = mydataset %>% select(1, 4, 5, 6) %>% distinct() %>% filter(chr != "chrY", chr != "chrM")

data = intron_data %>% left_join(trans_info,  by = "trans_id") %>% left_join(convert, by = c("gene_id", "trans_id")) %>% select(4, 2, 3, 5, 1, 6, 7, 8, 9, 10)

iso_intron_data = as.data.table(data)

save(iso_intron_data, file = "~/isoviz/data/iso_intron_data.rda")

```

```{r}
load(file = "~/isoviz/data/iso_intron_data.rda")
load(file = "~/isoviz/data/leafcutter_clusters.rda")

# map junctions function- for now, it isn't using cell type but eventually we should change this
map_junction = function(gene = "ENSG00000001167.15", cell_type = "A375"){

  # filter for gene, for now, also filter for protein coding transcripts- can set this as a parameter later
  gene_iso_data = iso_intron_data %>% filter(grepl(gene, gene_id), transcript_type == "protein_coding")

  # join iso data with leafcutter data and mark any unannotated junctions
  gene_cluster = gene_iso_data %>% left_join(leafcutter_clusters, by = c("chr", "intron_starts" = "start", "intron_ends" = "end")) %>% filter(!is.na(cluster.n))
  gene = unique(gene_cluster$gene_id)
  name = unique(gene_cluster$gene_name)
  gene_strand = unique(gene_cluster$strand)
  cluster_ids = unique(gene_cluster$cluster.n)
  output = leafcutter_clusters %>% filter(cluster.n %in% cluster_ids) %>% group_by(cluster.n) %>% mutate(junc.per.cluster = n()) %>% 
    left_join(gene_iso_data, by = c("chr", "start" = "intron_starts", "end" = "intron_ends")) %>% 
    mutate(junc.usage = round((junc.counts/cluster.counts)*100, digits = 0), 
           transcript_name = ifelse(is.na(transcript_name), "Unannotated", transcript_name),
           trans_id = ifelse(is.na(trans_id), "Unannotated", trans_id),
           gene_id = ifelse(is.na(gene_id), gene, gene_id),
           gene_name = ifelse(is.na(gene_name), name, gene_name),
           strand = ifelse(is.na(strand), gene_strand, strand))

  return(output)
}

to_plot = map_junction("ENSG00000001167.15") # NFYA
to_plot = map_junction("ENSG00000100320") # RBFOX2

```

```{r}
load(file='~/isoviz/data/iso_data_complete.rda')
to_plot

# plot isoform plus leafcutter clusters
plot_junction_to_isoform = function(to_plot){

  # get info
  gene = unique(to_plot$gene_id)
  name = unique(to_plot$gene_name)
  transcripts = unique(to_plot$transcript_name)

  # I am copying the plot_isoforms function here, figure out how to integrate these later
  gene_trans =which(str_detect(iso_data_complete$gene_id, gene))
  gene_data = iso_data_complete[gene_trans,]

  # filter for transcripts that have junction level data to support
  gene_data %<>% filter(transcript_name %in% transcripts)

  # setting parameters for the plot based on length of the gene and number of isoforms
  length = max(gene_data$end) - min(gene_data$start)
  min_start = min(gene_data$start)
  n = gene_data %>% group_by(trans_id) %>% tally() %>% nrow()
  y_max = n + 1
  max_end = max(gene_data$end)

  # order the isoforms by length for plotting
  order_plot = gene_data %>% group_by(transcript_name, transcript_length) %>% tally() %>%
    dplyr::arrange(desc(transcript_length), desc(n))
  order_plot$trans_order = 1:nrow(order_plot)
  order_plot = order_plot %>% dplyr::select(transcript_name, trans_order)

  to_plot_ordered = gene_data %>% full_join(order_plot, by = c("transcript_name")) %>%
    arrange(trans_order, blockstarts) %>%
    mutate(y1 = trans_order - 0.2, y2 = trans_order + 0.2)
  to_plot_ordered = to_plot_ordered %>% mutate(seg_end = end)

  # make isoform plot
  p1 = ggplot() +
    geom_segment(aes(x = to_plot_ordered$start, y = to_plot_ordered$trans_order,
                     xend = to_plot_ordered$seg_end, yend = to_plot_ordered$trans_order)) +
    geom_rect(data = to_plot_ordered, mapping = aes(xmin = blockstarts, xmax = blockends, ymin = y1, ymax = y2)) +
    xlim(min_start, max_end) + ylim(0,y_max) +
    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$transcript_name)) +
    theme_bw() + ylab("") + xlab("") +
    theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "top") +
    ggtitle(paste0(name, " (", to_plot_ordered$strand[1], ")")) +
    theme(axis.text = element_text(size = 10), plot.title = element_text(hjust = 0.5)) + coord_cartesian(xlim = c(min_start, max_end))
    #coord_cartesian(clip = 'off')
    #theme(plot.margin = margin(0.5,1.5,0.5,0.5, "in"))

  # new code to plot leafcutter cluster info underneath
  clusters = unique(to_plot$cluster.n)
  introns = to_plot %>% select(-gene_id, -trans_id, -gene_name, -transcript_name) %>% distinct() %>% group_by(cluster.n) %>% mutate(junc_order = row_number())
  p2 = ggplot() + geom_rect(data = introns, mapping = aes(xmin = start, xmax = end, ymin = junc_order - 0.2, ymax = junc_order + 0.2)) + 
    xlim(min_start, max_end) + theme_bw() + facet_wrap(~ cluster.n, scales = "free_y", ncol = 1, strip.position = "top") + 
    xlab("Hg38 Genomic Position (bp)") + ylab("") + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
    coord_cartesian(xlim = c(min_start, max_end))

  plot_grid(p1, p2, ncol = 1, align = "v", rel_heights = c(3, 3))
  ggsave("~/isoviz/plots/rbfox2_isoform_draft.pdf", width = 8, height = 4)

    # try using pivot for aggregating isoform

}
```

### 2023-03-28 temp_code/MS_isoviz_code.Rmd
MS add plot junction to isoform

@@
 
+Raw data -> Data for package
+Leafcutter clusters
+Run this on publicly available data for major cell types
+But have function for people to read in their on bam file
+
+Defining Intron clustering with Leafcutter
+make a text file with the name of all bed files
+
+# more stringent- better for calling unique clusters
+python ~/Downloads/leafcutter/clustering/leafcutter_cluster_regtools.py -j a375_screen_cells_for_lfc.txt -m 30 -p 0.1 -o a375_screen_cells -l 500000 --includeconst --checkchrom
+
+# less stringent (-m 5) which is better for detecting all singletons
+python ~/Downloads/leafcutter/clustering/leafcutter_cluster_regtools.py -j a375_screen_cells_for_lfc.txt -m 5 -p 0.1 -o a375_screen_cells_5reads -l 500000 --includeconst --checkchrom
+
+output is a text file
+-p MINCLURATIO, --mincluratio=MINCLURATIO
+                        minimum fraction of reads in a cluster that support a
+                        junction (default 0.001)
+
+--includeconst
 ```{r}
-# generate leafcutter cluster files and store for standard cell types
-# this is publicly available A375 data used to design the screen
-
-# get leafcutter data in proper format
-clusters <- read_tsv("~/isoviz/data-raw/SRR6515912_out-regtools_.junc.a375_m30_p0.01_.sorted", col_names = FALSE, skip = 1) # 15,901
+# This is data from our A375 screen cells and is the most accurate for our work
+# now this includes constitutive exons which have a single junction per cluster
+# required 30 total reads per cluster and junction to be 10% of total
+# total reads in a375-cyt_Aligned.q20.sort.bam: 43,407,715
+clusters <- read_tsv("~/isoviz/data-raw/a375_screen_cells_perind.constcounts.gz", col_names = FALSE, skip = 1) # 59,783
 clusters %<>% 
@@
 
-# remove clusters with no counts, not sure why these are here to begin with- only 7
+# remove clusters with no counts
 # filter out chrM and chrY clusters
@@
 
-leafcutter_clusters = as.data.table(clusters)
+a375_screen_lfc_clusters = as.data.table(clusters)
+save(a375_screen_lfc_clusters, file = "~/isoviz/data/a375_screen_lfc_clusters.rda")
 
-save(leafcutter_clusters, file = "~/isoviz/data/leafcutter_clusters.rda")
 ```
 
+Raw data -> Data for package
+Gene/transcript name conversion file
+Others might want to do this for a different gtf version
 ```{r}
@@
 # create gene name conversion file from gtf
-gtf = read_tsv("~/isoviz/data-raw/gencode.v41.basic.annotation.gtf", col_names = FALSE, skip = 5) %>% filter(X3 == "transcript")
+
+gtf_convert = function(gtf_file){
+  gtf = read_tsv(gtf_file, col_names = FALSE, skip = 5) %>% filter(X3 == "transcript")
+
   gtf$transcript_id = sapply(strsplit(as.character(gtf$X9), ";"), function(x) x[[2]])
@@
 
-gtf %>% dplyr::select(gene_id, trans_id = transcript_id, gene_name, transcript_name, gene_type, transcript_type) %>% distinct() %>%
-  write_tsv("~/isoviz/data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)
+  gtf %<>% dplyr::select(gene_id, trans_id = transcript_id, gene_name, transcript_name, gene_type, transcript_type) %>%
+    distinct()
+
+}
+
+output = gtf_convert("~/isoviz/data-raw/gencode.v41.basic.annotation.gtf")
+write_tsv(output, "~/isoviz/data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)
 
@@
 
+Add this to original genomedata.R script
 ```{r}
@@
 
+Write function to generate intron coords from exon coords- this is currently slow (~30 minutes), can we make the loop more efficient?
 ```{r}
@@
 
-mydataset <- fread("~/isoviz/data-raw/gencode.v41.basic.annotation.psl")
+make_intron_coords = function(psl){
+
+  mydataset <- fread(psl)
 
@@
 
+  # this is the part that is slow
   for(id in transcript_ids){
@@
 
-data = intron_data %>% left_join(trans_info,  by = "trans_id") %>% left_join(convert, by = c("gene_id", "trans_id")) %>% select(4, 2, 3, 5, 1, 6, 7, 8, 9, 10)
+  data = intron_data %>% left_join(trans_info,  by = "trans_id") %>% 
+    left_join(convert, by = c("gene_id", "trans_id")) %>% select(4, 2, 3, 5, 1, 6, 7, 8, 9, 10)
 
@@
 
+}
+
+iso_intron_data = make_intron_coords("~/isoviz/data-raw/gencode.v41.basic.annotation.psl")
 save(iso_intron_data, file = "~/isoviz/data/iso_intron_data.rda")
@@
 
+Function: Map junction to isoform
 ```{r}
 load(file = "~/isoviz/data/iso_intron_data.rda")
-load(file = "~/isoviz/data/leafcutter_clusters.rda")
+load(file = "~/isoviz/data/a375_screen_lfc_clusters.rda")
 
@@
   # filter for gene, for now, also filter for protein coding transcripts- can set this as a parameter later
-  gene_iso_data = iso_intron_data %>% filter(grepl(gene, gene_id), transcript_type == "protein_coding")
+  gene_iso_data = iso_intron_data %>% filter(grepl(gene, gene_id), transcript_type == "protein_coding") # 17 for NFYA
 
   # join iso data with leafcutter data and mark any unannotated junctions
-  gene_cluster = gene_iso_data %>% left_join(leafcutter_clusters, by = c("chr", "intron_starts" = "start", "intron_ends" = "end")) %>% filter(!is.na(cluster.n))
+  const_junc_list = a375_screen_lfc_clusters %>% 
+    group_by(cluster.n) %>% 
+    tally() %>% filter(n == 1) %>% ungroup()
+  a375_screen_lfc_clusters %<>% mutate(junc.type = ifelse(cluster.n %in% const_junc_list$cluster.n, "Common", "Unique"))
+  gene_cluster = gene_iso_data %>% 
+    left_join(a375_screen_lfc_clusters, by = c("chr", "intron_starts" = "start", "intron_ends" = "end")) %>%
+    filter(!is.na(cluster.n))
+
   gene = unique(gene_cluster$gene_id)
@@
   cluster_ids = unique(gene_cluster$cluster.n)
-  output = leafcutter_clusters %>% filter(cluster.n %in% cluster_ids) %>% group_by(cluster.n) %>% mutate(junc.per.cluster = n()) %>% 
+
+  output = a375_screen_lfc_clusters %>% filter(cluster.n %in% cluster_ids) %>% group_by(cluster.n) %>% mutate(junc.per.cluster = n()) %>% 
     left_join(gene_iso_data, by = c("chr", "start" = "intron_starts", "end" = "intron_ends")) %>% 
     mutate(junc.usage = round((junc.counts/cluster.counts)*100, digits = 0), 
-           transcript_name = ifelse(is.na(transcript_name), "Unannotated", transcript_name),
-           trans_id = ifelse(is.na(trans_id), "Unannotated", trans_id),
+           transcript_name = ifelse(is.na(transcript_name), paste0(cluster.n, "_Unannotated"), transcript_name),
+           trans_id = ifelse(is.na(trans_id), paste0(cluster.n, "_Unannotated"), trans_id),
            gene_id = ifelse(is.na(gene_id), gene, gene_id),
@@
 to_plot = map_junction("ENSG00000001167.15") # NFYA
-to_plot = map_junction("ENSG00000100320") # RBFOX2
+to_plot = map_junction("ENSG00000100320") # RBFOX2, alt start is missing because it's intron doesnt overlap another and counts are only 8 so included in consitutive either
+
+# can we ask if all junctions of an isoform are represented to filter isoforms?
+# can we create pseudo regtools output based on all possible junctions and run that through leafcutter to get true common versus unique
 
@@
 load(file='~/isoviz/data/iso_data_complete.rda')
-to_plot
+df = to_plot
 
 # plot isoform plus leafcutter clusters
-plot_junction_to_isoform = function(to_plot){
+plot_junction_to_isoform = function(df, include_common_juncs = FALSE){
+
+  # I am copying the plot_isoforms function here, figure out how to integrate these later
+  gene = unique(df$gene_id)
+  gene_trans =which(str_detect(iso_data_complete$gene_id, gene))
+  gene_data = iso_data_complete[gene_trans,]
+
+  # filter out common junctions for plotting (default)
+  if(include_common_juncs == FALSE){
+    df %<>% filter(junc.type == "Unique")
+  }
 
   # get info
-  gene = unique(to_plot$gene_id)
-  name = unique(to_plot$gene_name)
-  transcripts = unique(to_plot$transcript_name)
-
-  # I am copying the plot_isoforms function here, figure out how to integrate these later
-  gene_trans =which(str_detect(iso_data_complete$gene_id, gene))
-  gene_data = iso_data_complete[gene_trans,]
+  name = unique(df$gene_name)
+  transcripts = unique(df$transcript_name)
 
@@
   to_plot_ordered = gene_data %>% full_join(order_plot, by = c("transcript_name")) %>%
-    arrange(trans_order, blockstarts) %>%
-    mutate(y1 = trans_order - 0.2, y2 = trans_order + 0.2)
+    arrange(trans_order, blockstarts) 
   to_plot_ordered = to_plot_ordered %>% mutate(seg_end = end)
@@
                      xend = to_plot_ordered$seg_end, yend = to_plot_ordered$trans_order)) +
-    geom_rect(data = to_plot_ordered, mapping = aes(xmin = blockstarts, xmax = blockends, ymin = y1, ymax = y2)) +
+    geom_rect(data = to_plot_ordered, mapping = aes(xmin = blockstarts, xmax = blockends, ymin = trans_order - 0.3, ymax = trans_order + 0.3)) +
     xlim(min_start, max_end) + ylim(0,y_max) +
-    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$transcript_name)) +
+    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$transcript_name), size = 3) +
     theme_bw() + ylab("") + xlab("") +
-    theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "top") +
-    ggtitle(paste0(name, " (", to_plot_ordered$strand[1], ")")) +
-    theme(axis.text = element_text(size = 10), plot.title = element_text(hjust = 0.5)) + coord_cartesian(xlim = c(min_start, max_end))
-    #coord_cartesian(clip = 'off')
-    #theme(plot.margin = margin(0.5,1.5,0.5,0.5, "in"))
+    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), 
+          axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "top", 
+          plot.title = element_text(hjust = 0.5), plot.margin = margin(0.1,1,0,0.1, "in")) +
+    ggtitle(paste0(name, " Junction to Isoform Map (" , to_plot_ordered$strand[1], " strand )")) + 
+    coord_cartesian(xlim = c(min_start, max_end), clip = 'off')
 
-  # new code to plot leafcutter cluster info underneath
-  clusters = unique(to_plot$cluster.n)
-  introns = to_plot %>% select(-gene_id, -trans_id, -gene_name, -transcript_name) %>% distinct() %>% group_by(cluster.n) %>% mutate(junc_order = row_number())
-  p2 = ggplot() + geom_rect(data = introns, mapping = aes(xmin = start, xmax = end, ymin = junc_order - 0.2, ymax = junc_order + 0.2)) + 
-    xlim(min_start, max_end) + theme_bw() + facet_wrap(~ cluster.n, scales = "free_y", ncol = 1, strip.position = "top") + 
-    xlab("Hg38 Genomic Position (bp)") + ylab("") + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
+  # new code to plot leafcutter cluster info underneath isoform level data
+  # aggregate isoform information
+  text = df %>% separate(col = transcript_name, into = c("name", "transcript")) %>% group_by(cluster.n, start, end) %>% 
+    arrange(cluster.n, start, end, transcript) %>%
+    mutate(isoforms = paste0(transcript, collapse = ",")) %>%
+    select(1:5, junc.usage, isoforms) %>% distinct() %>% ungroup() %>% 
+    mutate(text_plot = paste0(junc.counts, " reads (", junc.usage, "%) : ", isoforms))
+
+  clusters = unique(df$cluster.n)
+  introns = df %>% select(-gene_id, -trans_id, -gene_name, -transcript_name) %>% 
+    distinct() %>% arrange(cluster.n, start)
+  introns$junc.order = 1:nrow(introns)
+  introns %<>% left_join(text)
+  p2 = ggplot() + 
+    geom_rect(data = introns, mapping = aes(xmin = start, xmax = end, ymin = junc.order - 0.2, ymax = junc.order + 0.2, fill = cluster.n)) + 
+    geom_text(data = introns, aes(x = end, y = junc.order, label = text_plot), hjust = 0, size = 3) +
+    xlim(min_start, max_end) + theme_bw() + 
+    xlab("Hg38 Genomic Position (bp)") + ylab("") + 
+    theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "none", plot.margin = margin(0,1,0,0.1, "in")) +
     coord_cartesian(xlim = c(min_start, max_end))
 
-  plot_grid(p1, p2, ncol = 1, align = "v", rel_heights = c(3, 3))
-  ggsave("~/isoviz/plots/rbfox2_isoform_draft.pdf", width = 8, height = 4)
+  # plot dimenstions based on number of isoforms and junctions
+  t = (n + nrow(introns))
+  h = t/2
+  p1_p2 = plot_grid(p1, p2, ncol = 1, align = "v", rel_heights = c(n*0.8, nrow(introns)))
+  #ggsave("~/isoviz/plots/rbfox2_isoforms.pdf", width = 8, height = h)
+  return(p1_p2)
+}
 
-    # try using pivot for aggregating isoform
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/rbfox2_isoforms.pdf", width = 8, height = 8.5)
+```
 
-}
+```{r}
+to_plot = map_junction("ENSG00000001167.15") # NFYA
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/nfya_isoforms.pdf", width = 8, height = 4)
+
+to_plot = map_junction("ENSG00000100320") # RBFOX2, alt start is missing because it's intron doesnt overlap another and counts are only 8 so included in consitutive either
+plot_junction_to_isoform(to_plot)
 ```

### 2023-04-04 temp_code/MS_isoviz_code.Rmd
Updated MS_isoviz_code

@@
 - could we even have a screen design setting? to scale it up?
+- also, can use the code and plots to quantify knockdown
+- need to control for library size- tpm of gene?
 
@@
 # required 30 total reads per cluster and junction to be 10% of total
-# total reads in a375-cyt_Aligned.q20.sort.bam: 43,407,715
 clusters <- read_tsv("~/isoviz/data-raw/a375_screen_cells_perind.constcounts.gz", col_names = FALSE, skip = 1) # 59,783
@@
 
-a375_screen_lfc_clusters = as.data.table(clusters)
-save(a375_screen_lfc_clusters, file = "~/isoviz/data/a375_screen_lfc_clusters.rda")
+clusters_table = as.data.table(clusters)
+save(clusters_table, file = "~/isoviz/data/A375_lfc_clusters.rda")
+
+# HEK293 cells- our data
+clusters <- read_tsv("~/isoviz/data-raw/HEK-NFYA-rep2-empty_perind.constcounts.gz", col_names = FALSE, skip = 1) # 46,102
+clusters %<>% 
+  separate(X1, into = c("coords", "counts"), sep = " ") %>% 
+  separate(counts, into = c("junc.counts", "cluster.counts"), sep = "/", convert = TRUE) %>%
+  separate(coords, into = c("chr", "start", "end", "name"), sep = ":", convert = TRUE) %>% 
+  separate(name, into = c("same", "cluster.n", "extra"), sep = "_") %>% 
+  select(-same, -extra)
+
+# remove clusters with no counts
+# filter out chrM and chrY clusters
+clusters %<>% filter(cluster.counts > 0, chr != "chrY", chr != "chrM")
+clusters_table = as.data.table(clusters)
+save(clusters_table, file = "~/isoviz/data/HEK293_lfc_clusters.rda")
+
+# 35738736-36028824
+singletons <- read_tsv("~/knowles_lab/Megan/230320_cas13_validation/junctions/HEK-NFYA-rep2-empty_singletons_perind.constcounts.gz", col_names = FALSE, skip = 1)
+singletons %<>% 
+  separate(X1, into = c("coords", "counts"), sep = " ") %>% 
+  separate(counts, into = c("junc.counts", "cluster.counts"), sep = "/", convert = TRUE) %>%
+  separate(coords, into = c("chr", "start", "end", "name"), sep = ":", convert = TRUE) %>% 
+  separate(name, into = c("same", "cluster.n", "extra"), sep = "_") %>% 
+  select(-same, -extra)
+
+singletons %>% filter(chr == "chr22", start >= 35738736, end <= 36028824)
+
+# hESC- our data
+clusters <- read_tsv("~/isoviz/data-raw/21c_hESC_RB2_empty_perind.constcounts.gz", col_names = FALSE, skip = 1)
+clusters %<>% 
+  separate(X1, into = c("coords", "counts"), sep = " ") %>% 
+  separate(counts, into = c("junc.counts", "cluster.counts"), sep = "/", convert = TRUE) %>%
+  separate(coords, into = c("chr", "start", "end", "name"), sep = ":", convert = TRUE) %>% 
+  separate(name, into = c("same", "cluster.n", "extra"), sep = "_") %>% 
+  select(-same, -extra)
+
+# remove clusters with no counts
+# filter out chrM and chrY clusters
+clusters %<>% filter(cluster.counts > 0, chr != "chrY", chr != "chrM")
+clusters_table = as.data.table(clusters)
+save(clusters_table, file = "~/isoviz/data/hESC_lfc_clusters.rda")
 
@@
 load(file = "~/isoviz/data/iso_intron_data.rda")
-load(file = "~/isoviz/data/a375_screen_lfc_clusters.rda")
+#load(file = "~/isoviz/data/a375_screen_lfc_clusters.rda")
+#load(file = "~/isoviz/data/HEK293_lfc_clusters.rda")
 
+# is it better to load rda within a function?
 # map junctions function- for now, it isn't using cell type but eventually we should change this
-map_junction = function(gene = "ENSG00000001167.15", cell_type = "A375"){
+map_junction = function(gene = "ENSG00000001167.15", cell_type = "HEK293"){
 
@@
   # join iso data with leafcutter data and mark any unannotated junctions
-  const_junc_list = a375_screen_lfc_clusters %>% 
+  # consider cell type
+  lfc_clusters = paste0(cell_type, "_lfc_clusters")
+  load(file = paste0("~/isoviz/data/", lfc_clusters, ".rda"))
+  const_junc_list = clusters_table %>% 
     group_by(cluster.n) %>% 
     tally() %>% filter(n == 1) %>% ungroup()
-  a375_screen_lfc_clusters %<>% mutate(junc.type = ifelse(cluster.n %in% const_junc_list$cluster.n, "Common", "Unique"))
+  clusters_table %<>% mutate(junc.type = ifelse(cluster.n %in% const_junc_list$cluster.n, "Common", "Unique"))
   gene_cluster = gene_iso_data %>% 
-    left_join(a375_screen_lfc_clusters, by = c("chr", "intron_starts" = "start", "intron_ends" = "end")) %>%
+    left_join(clusters_table, by = c("chr", "intron_starts" = "start", "intron_ends" = "end")) %>%
     filter(!is.na(cluster.n))
@@
 
-  output = a375_screen_lfc_clusters %>% filter(cluster.n %in% cluster_ids) %>% group_by(cluster.n) %>% mutate(junc.per.cluster = n()) %>% 
+  output = clusters_table %>% filter(cluster.n %in% cluster_ids) %>% group_by(cluster.n) %>% mutate(junc.per.cluster = n()) %>% 
     left_join(gene_iso_data, by = c("chr", "start" = "intron_starts", "end" = "intron_ends")) %>% 
@@
            gene_name = ifelse(is.na(gene_name), name, gene_name),
-           strand = ifelse(is.na(strand), gene_strand, strand))
+           strand = ifelse(is.na(strand), gene_strand, strand), cell_line = paste0(cell_type))
 
@@
 ```{r}
-to_plot = map_junction("ENSG00000001167.15") # NFYA
+# NFYA
+to_plot = map_junction("ENSG00000001167.15", cell_type = "A375") 
 plot_junction_to_isoform(to_plot)
@@
 
-to_plot = map_junction("ENSG00000100320") # RBFOX2, alt start is missing because it's intron doesnt overlap another and counts are only 8 so included in consitutive either
+to_plot = map_junction("ENSG00000001167.15", cell_type = "HEK293") 
 plot_junction_to_isoform(to_plot)
+# for height, count rows and divide by 2
+ggsave("~/isoviz/plots/nfya_isoforms_hek293.pdf", width = 8, height = 2)
+
+to_plot = map_junction("ENSG00000001167.15", cell_type = "hESC") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/nfya_isoforms_hESC.pdf", width = 8, height = 2)
+
+# RBFOX2
+to_plot = map_junction("ENSG00000100320", cell_type = "A375") # alt start is missing because it's intron doesnt overlap another and counts are only 8 so included in consitutive either
+plot_junction_to_isoform(to_plot)
+
+to_plot = map_junction("ENSG00000100320", cell_type = "HEK293") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/rbfox2_isoforms_hek293.pdf", width = 8, height = 9)
+
+to_plot = map_junction("ENSG00000100320", cell_type = "hESC") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/rbfox2_isoforms_hESC.pdf", width = 8, height = 8.5)
+
 ```
+
+Generate a junction file of all possible junctions sequences for Andrew to make predictions
+- use gencode psl
+- want to make sure to have junc id that will correspond to leafcutter (can merge by sequence at first)
+- can make this information into a table
+```{r}
+load(file = "~/isoviz/data/iso_intron_data.rda")
+iso_intron_data %>% group_by(transcript_type) %>% tally()
+iso_intron_data %>% group_by(chr) %>% tally()
+iso_intron_data %>% filter(is.na(chr))
+
+# filter for protein coding and lncRNAs
+iso_intron_pc = iso_intron_data %>% filter(transcript_type == "protein_coding" | transcript_type == "lncRNA")
+
+# remove transcript ids and collapse on junction coords
+collapsed_junctions = iso_intron_pc %>% select(-trans_id, -transcript_name, -gene_type, -transcript_type) %>% 
+  distinct() %>% select(chr, start = intron_starts, end = intron_ends, gene_id, gene_name, strand) # 281,880 junctions- how many show any uniqueness?
+
+# to address how many annotated isoforms have a truly unique junction
+# first filter the genes with more than one transcript
+iso_intron_pc %>% select(gene_id, trans_id) %>% distinct() # 85,586 transcripts
+n_df = iso_intron_pc %>% select(gene_id, trans_id) %>% distinct() %>% group_by(gene_id) %>% mutate(n_transcripts = n()) %>% ungroup() # 33,413
+gene_list = n_df %>% select(-trans_id) %>% distinct() %>% filter(n_transcripts > 1) %>% select(gene_id) # 17,035 genes have more than 1 isoform
+
+iso_pc_df = iso_intron_pc %>% full_join(n_df) %>% filter(gene_id %in% gene_list$gene_id) %>%
+  group_by(chr, intron_starts, intron_ends, gene_id, strand, gene_name, n_transcripts) %>% mutate(n_trans_per_junc = n(), commonality = round((n_trans_per_junc/n_transcripts)*100, 0)) %>%
+  mutate(junction_category = ifelse(n_trans_per_junc == 1, "fully_unique", ifelse(commonality == 100, "common", "partial_unique"))) %>% ungroup()
+
+# junction level numbers- there are ~50,000 fully unique junctions for pc genes, meaning they belong to a singular isoform
+iso_pc_df %>% select(chr, intron_starts, intron_ends, strand, gene_id, gene_name, gene_type, junction_category) %>% distinct() %>% 
+  group_by(gene_type, junction_category) %>% tally()
+
+# isoform-level, how many isoforms are uniquely targetable
+iso_pc_df %>% filter(junction_category == "fully_unique") %>% select(gene_id, gene_name, trans_id, transcript_name, gene_type) %>% distinct() # 47,657 transcripts are uniquely targetable
+iso_pc_df %>% filter(junction_category == "fully_unique") %>% select(gene_id, gene_name, trans_id, transcript_name, gene_type) %>% distinct() %>%
+  distinct(gene_name) # 16,801 genes
+
+```

### 2023-04-04 temp_code/leafcutter_cluster_regtools.py

@@ -1,0 +1,538 @@
+#!/usr/bin/env python
+
+# https://github.com/griffithlab/regtools
+# /home/yangili1/tools/regtools/build/regtools junctions extract -a 8 -i 50 -I 500000 bamfile.bam -o outfile.junc
+# Using regtools speeds up the junction extraction step by an order of magnitude or more
+
+import sys
+import tempfile
+import os
+import gzip
+import shutil
+
+def main(options,libl):
+
+    if options.cluster is None:
+        pool_junc_reads(libl, options)
+        refine_clusters(options)
+
+    sort_junctions(libl, options)
+    merge_junctions(options)
+    get_numers(options)
+
+def pool_junc_reads(flist, options):
+
+    outPrefix = options.outprefix
+    rundir = options.rundir
+    maxIntronLen = int(options.maxintronlen)
+    nochromcheck = options.nochromcheck
+
+    outFile = "%s/%s_pooled"%(rundir,outPrefix)
+
+    chromLst = ["chr%d"%x for x in range(1,23)]+['chrX','chrY']+["%d"%x for x in range(1,23)]+['X','Y']
+    by_chrom = {}
+
+    # check that all files exist
+    for libl in flist:
+        if not os.path.isfile(libl.strip()):
+            raise ValueError("File %s does not exist."%lib)
+
+    for k,libl in enumerate(flist, 1):
+        lib = libl.strip()
+        if options.verbose:
+            sys.stderr.write("scanning %d/%d: %s ...\n"%(k, len(flist), lib))
+
+        if lib.endswith('.gz'):
+            opener = gzip.open(lib, 'rt')
+        else:
+            opener = open(lib, 'r')
+
+        for ln in opener:
+            lnsplit = ln.split()
+            if len(lnsplit) < 6:
+                sys.stderr.write("Error in %s\n" % lib)
+                continue
+            chrom, A, B, dot, counts, strand, rA, rb, rgb, blockCount, blockSize, blockStarts = lnsplit
+            if int(blockCount) > 2:
+                print(ln)
+                continue
+            # regtools -s 0 (unstranded) now puts "?" in strand field when strand is ambiguous
+            if strand == "?": continue
+            if not nochromcheck and (chrom not in chromLst): continue  # excludes ALT contigs etc
+            Aoff, Boff = blockSize.split(",")
+            A, B = int(A)+int(Aoff), int(B)-int(Boff)+1
+
+            if B-A > int(maxIntronLen): continue
[…]

### 2023-04-20 temp_code/MS_isoviz_code.Rmd
update Megan's code

@@
 require(cowplot)
+require(Biostrings)
+require(rempsyc) # for fancy tables: https://cran.r-project.org/web/packages/rempsyc/vignettes/table.html
 
@@
 
+module load python/2.7.8
+
+/gpfs/commons/home/mschertzer/knowles_lab/Megan/230320_cas13_validation/junctions
+
 # more stringent- better for calling unique clusters
@@
 
+# moderate stringency
+python ~/Downloads/leafcutter/clustering/leafcutter_cluster_regtools.py -j 21c_hESC_RB2_empty_for_lfc.txt -m 20 -p 0.01 -o 21c_hESC_RB2_empty_p.01 -l 500000 --includeconst --checkchrom
+
 # less stringent (-m 5) which is better for detecting all singletons
 python ~/Downloads/leafcutter/clustering/leafcutter_cluster_regtools.py -j a375_screen_cells_for_lfc.txt -m 5 -p 0.1 -o a375_screen_cells_5reads -l 500000 --includeconst --checkchrom
+
+python ~/Downloads/leafcutter/clustering/leafcutter_cluster_regtools.py -j 21c_hESC_RB2_empty_for_lfc.txt -m 1 -p 0.01 -o hesc_all_junctions_lfc -l 500000 --includeconst --checkchrom
 
@@
                         junction (default 0.001)
+-m MINCLUREADS, --minclureads=MINCLUREADS
+                        minimum reads in a cluster (default 30 reads)                       
+--includeconst
+            also include constitutive introns
 
---includeconst
+Inputs leafcutter clustering output and outputs formatted table
 ```{r}
@@
 
-# remove clusters with no counts
-# filter out chrM and chrY clusters
 clusters %<>% filter(cluster.counts > 0, chr != "chrY", chr != "chrM")
@@
 save(clusters_table, file = "~/isoviz/data/hESC_lfc_clusters.rda")
+
+# hESC- medium stringent
+clusters <- read_tsv("~/isoviz/data-raw/21c_hESC_RB2_empty_p.01_perind.constcounts.gz", col_names = FALSE, skip = 1)
+clusters %<>% 
+  separate(X1, into = c("coords", "counts"), sep = " ") %>% 
+  separate(counts, into = c("junc.counts", "cluster.counts"), sep = "/", convert = TRUE) %>%
+  separate(coords, into = c("chr", "start", "end", "name"), sep = ":", convert = TRUE) %>% 
+  separate(name, into = c("same", "cluster.n", "extra"), sep = "_") %>% 
+  select(-same, -extra)
+
+clusters %<>% filter(cluster.counts > 0, chr != "chrY", chr != "chrM")
+clusters_table = as.data.table(clusters)
+save(clusters_table, file = "~/isoviz/data/hESC_med_string_lfc_clusters.rda")
+
+```
+
+# Consider all junction counts and use gencode annotations to define unique versus common
+```{r}
+clusters <- read_tsv("~/isoviz/data-raw/hesc_all_junctions_lfc_perind.constcounts.gz", col_names = FALSE, skip = 1)
+clusters %<>% 
+  separate(X1, into = c("coords", "counts"), sep = " ") %>% 
+  separate(counts, into = c("junc.counts", "cluster.counts"), sep = "/", convert = TRUE) %>%
+  separate(coords, into = c("chr", "start", "end", "name"), sep = ":", convert = TRUE) %>% 
+  separate(name, into = c("same", "cluster.n", "extra"), sep = "_") %>% 
+  select(-same, -extra)
+
+clusters %<>% filter(cluster.counts > 0, chr != "chrY", chr != "chrM")
+clusters_table = as.data.table(clusters)
+save(clusters_table, file = "~/isoviz/data/hESC_all_lfc_junctions.rda")
 
@@
 Write function to generate intron coords from exon coords- this is currently slow (~30 minutes), can we make the loop more efficient?
+This is only done for the gencode gtf for now which would be provided, this chunk would only need to run if people had their own gtf they wanted to use
 ```{r}
@@
 
-to_plot = map_junction("ENSG00000001167.15") # NFYA
-to_plot = map_junction("ENSG00000100320") # RBFOX2, alt start is missing because it's intron doesnt overlap another and counts are only 8 so included in consitutive either
+#to_plot = map_junction("ENSG00000001167.15") # NFYA
+#to_plot = map_junction("ENSG00000100320") # RBFOX2, alt start is missing because it's intron doesnt overlap another and counts are only 8 so included in consitutive either
 
@@
 
+Updated Function: Map junction to isoform: Updated to include all junctions and use gencode annotations to mark unique versus common
+How do we incorporate expression here since the hESC data is more deeply sequenced- could combine feature counts when we format the leafcutter output
+```{r}
+# merge with gencode info and predictions from 230411_cas13_paper_figures.Rmd
+# we should probably filter these for things like runs of T's, etc like we did when designing the screen
+# missing 166- which are guides in Harm's screen
+v41_info = read_tsv("~/isoviz/data/gencode_v41_all_junction_guides.txt", col_names = TRUE) %>% 
+  mutate(guide_sequence = as.character(guide_sequence)) # 2,255,040
+predictions = read_csv("/gpfs/commons/groups/knowles_lab/cas13_share/230414_predictions/combined-corrected-predictions-gencode.csv") %>% 
+  select(guide_id, predicted_lfc) # 2,254,874
+
+gencode_intron_all_data = v41_info %>% left_join(predictions, by = "guide_id")
+save(gencode_intron_all_data, file = "~/isoviz/data/gencode_intron_all_data.rda")
+
+# this gets read in within function: hESC_all_lfc_junctions.rda
+
+# map junctions function
+# I really don't need the guide info here just the junction info- consider revising and bring the guideRNAs in for another junction
+map_all_junctions = function(gene = "ENSG00000100320", cell_type = "hESC"){
+
+  # gencode- this is an updated file containing more junction info
+  load(file = "~/isoviz/data/gencode_intron_all_data.rda")
+
+  # filter for gene, for now, no need to filter for protein-coding since this gencode set only contains protein-coding
+  # this now contains all guide information for this gene
+  gene_iso_data = gencode_intron_all_data %>% filter(grepl(gene, gene_id))
+
+  # join iso data with leafcutter data and mark any unannotated junctions
+  # consider cell type
+  lfc_clusters = paste0(cell_type, "_all_lfc_junctions")
+  load(file = paste0("~/isoviz/data/", lfc_clusters, ".rda"))
+
+  # test
+  clusters_table %>% filter(chr == "chr22" & start > 35738735 & end < 36028824) %>% arrange(desc(start))
+
+  # since leafcutter junctions dont have gene annotations, need to combine here first
+  gene_cluster = gene_iso_data %>% 
+    left_join(clusters_table, by = c("chr", "junc_start" = "start", "junc_end" = "end"))
+
+  name = unique(gene_cluster$gene_name)
+  gene_strand = unique(gene_cluster$strand)
+
+  # to get information on leafcutter junctions that don't match annotation
+  all_gene_clusters = clusters_table %>% filter(cluster.n %in% gene_cluster$cluster.n) %>% 
+    left_join(gene_iso_data, by = c("chr", "start" = "junc_start", "end" = "junc_end")) %>%
+    filter(is.na(gene_id)) %>% 
+    mutate(gene_name = unique(gene_cluster$gene_name), strand = unique(gene_cluster$strand), 
+           gene_id = unique(gene_cluster$gene_id), junction_category = "unknown", 
+           transcript_isoforms = "unknown", junc_id = paste0("unk", 1:nrow(.))) %>%
+    select(chr, junc_start = start, junc_end = end, everything()) %>%
+    bind_rows(gene_cluster) %>%
+    mutate(junc.counts = ifelse(is.na(cluster.n), 0, junc.counts), cluster.counts = ifelse(is.na(cluster.n), 0, cluster.counts))
+
+  n = all_gene_clusters %>% distinct(cluster.n, junc_id) %>% group_by(cluster.n) %>% 
+    mutate(junc.per.cluster = ifelse(is.na(cluster.n), 1, n())) %>% ungroup()
+
+  all_gene_clusters %<>% left_join(n, by = c("cluster.n", "junc_id")) %>%
+    mutate(junc.usage = ifelse(junc.counts == 0, 0, round((junc.counts/cluster.counts)*100, digits = 0)), cell_line = paste0(cell_type))
+
+  #test
+  all_gene_clusters %>% filter(grepl("209", transcript_isoforms)) %>%
+    select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
+    distinct() %>% arrange(junc_start)
+
+  # create list of expressed isoforms
+  # the logic is that if an isoform has a fully unique junction with no counts, it is unlikely to be expressed
+  # I want to be careful here since some junctions won't be represented because of technical issues so only eliminate isoforms where none of their unique junctions have counts, avoid filtering on partially unique for now
+  non_expressed_genes = all_gene_clusters %>% select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
+    distinct() %>% separate_longer_delim(transcript_isoforms, delim = ",") %>% group_by(transcript_isoforms, junction_category) %>%
+    mutate(isoform_counts = ifelse(junction_category == "fully_unique", sum(junc.counts), junc.counts)) %>% 
+    filter(junction_category == "fully_unique" & isoform_counts == 0) %>% distinct(transcript_isoforms) %>% ungroup()
+
+  uniquely_targetable = all_gene_clusters %>% select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
+    distinct() %>% separate_longer_delim(transcript_isoforms, delim = ",") %>% filter(junction_category == "fully_unique") %>%
+    distinct(transcript_isoforms)
+
+  partially_targetable = all_gene_clusters %>% select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
+    distinct() %>% separate_longer_delim(transcript_isoforms, delim = ",") %>% filter(junction_category == "partial_unique") %>%
+    distinct(transcript_isoforms)
+
+  output = all_gene_clusters %>% select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
+    distinct() %>% separate_longer_delim(transcript_isoforms, delim = ",") %>%
+    mutate(transcript_targetable = ifelse(transcript_isoforms %in% uniquely_targetable$transcript_isoforms, "Uniquely", 
+                                          ifelse(transcript_isoforms %in% partially_targetable$transcript_isoforms, "Partially", "None"))) %>% 
+    mutate(isoform_expressed = ifelse(transcript_isoforms %in% non_expressed_genes$transcript_isoforms, "Unlikely", "Likely"))
+
+  if(nrow(output) == 0){
+    print(paste0("No junction reads detected for ", name, " in ", cell_type))
+  }
+
+  return(output)
+}
+
+test = map_all_junctions("ENSG00000100320", cell_type = "hESC")
+test %>% filter(isoform_expressed == "Likely") %>% distinct(transcript_isoforms)
+
+# chr22	35938897	36028240
+```
+
 ```{r}
 load(file='~/isoviz/data/iso_data_complete.rda')
-df = to_plot
+#df = to_plot
 
@@
   gene = unique(df$gene_id)
-  gene_trans =which(str_detect(iso_data_complete$gene_id, gene))
+  gene_trans = which(str_detect(iso_data_complete$gene_id, gene))
   gene_data = iso_data_complete[gene_trans,]
@@
 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/rbfox2_isoforms.pdf", width = 8, height = 8.5)
+#plot_junction_to_isoform(to_plot)
+#ggsave("~/isoviz/plots/rbfox2_isoforms.pdf", width = 8, height = 8.5)
+```
+
+Updated plotting to go with all junctions output
+```{r}
+# might have to use regtools if I truly want all junctions
+load(file='~/isoviz/data/iso_data_complete.rda')
+df = test
+
+# plot isoform plus leafcutter clusters
+plot_all_junction_to_isoform = function(df, junction_usage = 5, include_common_juncs = FALSE, filter_expressed_isoforms = TRUE){
+
+  # filter here based on input parameters
+  if(include_common_juncs == FALSE){
+    df %<>% filter(junction_category != "common")
+  }
+
+  if(filter_expressed_isoforms == TRUE){
+    df %<>% filter(isoform_expressed == "Likely")
+  }
+
+  to_remove = df %>% filter(junc.per.cluster == 2 & (junc.usage < junction_usage | junc.usage > (100 - junction_usage)))
+  df %<>% filter(!junc_id %in% to_remove$junc_id, junc.counts > 0)
+
+  # get info after filtering
+  name = unique(df$gene_name)
+  transcripts = unique(df$transcript_isoforms)
+
+  # filter transcripts for plotting too
+  gene_data = iso_data_complete %>% filter(transcript_name %in% df$transcript_isoforms)
+
+  # setting parameters for the plot based on length of the gene and number of isoforms
+  length = max(gene_data$end) - min(gene_data$start)
+  min_start = min(gene_data$start)
+  n = gene_data %>% group_by(trans_id) %>% tally() %>% nrow()
+  y_max = n + 1
+  max_end = max(gene_data$end)
+
+  # order the isoforms by length for plotting
+  order_plot = gene_data %>% group_by(transcript_name, transcript_length) %>% tally() %>%
+    dplyr::arrange(desc(transcript_length), desc(n))
+  order_plot$trans_order = 1:nrow(order_plot)
+  order_plot = order_plot %>% dplyr::select(transcript_name, trans_order)
+
+  to_plot_ordered = gene_data %>% full_join(order_plot, by = c("transcript_name")) %>%
+    arrange(trans_order, blockstarts) 
+  to_plot_ordered = to_plot_ordered %>% mutate(seg_end = end)
+
+  # make isoform plot
+  p1 = ggplot() +
+    geom_segment(aes(x = to_plot_ordered$start, y = to_plot_ordered$trans_order,
+                     xend = to_plot_ordered$seg_end, yend = to_plot_ordered$trans_order)) +
+    geom_rect(data = to_plot_ordered, mapping = aes(xmin = blockstarts, xmax = blockends, ymin = trans_order - 0.3, ymax = trans_order + 0.3)) +
+    xlim(min_start, max_end) + ylim(0,y_max) +
+    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$transcript_name), size = 3) +
+    theme_bw() + ylab("") + xlab("") +
+    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), 
+          axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "top", 
+          plot.title = element_text(hjust = 0.5), plot.margin = margin(0.1,1,0,0.1, "in")) +
+    ggtitle(paste0(name, " Junction to Isoform Map (" , to_plot_ordered$strand[1], " strand )")) + 
+    coord_cartesian(xlim = c(min_start, max_end), clip = 'off')
+
+  # new code to plot leafcutter cluster info underneath isoform level data
+  # aggregate isoform information
+  text = df %>% separate(col = transcript_isoforms, into = c("name", "transcript")) %>% 
+    mutate(transcript = ifelse(is.na(transcript), "unknown", transcript)) %>%
+    group_by(cluster.n, junc_id) %>% 
+    arrange(cluster.n, junc_id, transcript) %>%
+    mutate(isoforms = paste0(transcript, collapse = ",")) %>%
+    select(1:5, junc_id, junction_category, junc.usage, isoforms, junc.per.cluster) %>% distinct() %>% ungroup() %>% 
+    mutate(text_plot = paste0(junc.counts, " reads (", junc.usage, "%) : ", isoforms))
+
+  clusters = unique(df$cluster.n)
+  introns = df %>% select(-gene_id, -gene_name, -transcript_isoforms) %>% 
+    distinct() %>% arrange(cluster.n, junc_start)
+  introns$junc.order = 1:nrow(introns)
+  introns %<>% left_join(text)
+
+  # for guides - remove this or print out?
+  for_guides = introns %>% filter(grepl("junc", junc_id))
+
+  p2 = ggplot() + 
+    geom_rect(data = introns, mapping = aes(xmin = junc_start, xmax = junc_end, ymin = junc.order - 0.2, ymax = junc.order + 0.2, fill = cluster.n)) + 
+    geom_text(data = introns, aes(x = junc_end, y = junc.order, label = text_plot), hjust = 0, size = 3) +
+    xlim(min_start, max_end) + theme_bw() + 
+    xlab("Hg38 Genomic Position (bp)") + ylab("") + 
+    geom_text(data = introns, aes(x = Inf, y = junc.order, hjust = 0, label = junc_id), size = 3) + 
+    theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "none", plot.margin = margin(0,1,0,0.1, "in")) +
+    coord_cartesian(xlim = c(min_start, max_end), clip = 'off')
+
+  # plot dimenstions based on number of isoforms and junctions
+  t = (n + nrow(introns))
+  h = t/2
+  p1_p2 = plot_grid(p1, p2, ncol = 1, align = "v", rel_heights = c(n*0.8, nrow(introns)))
+
+  return(p1_p2)
+}
+
+ggsave("~/isoviz/plots/rbfox2_test.pdf", width = 8, height = h)
+
+#df %>% filter(transcript_isoforms == "RBFOX2-209") %>% arrange(junc_start)
+```
+
+Function to get guide RNA predictions for specific junctions
+```{r}
+# need to read in count information a different way- this is just a quick and dirty method
+# write a function to get expressed junctions/isoforms and call that within functions?
+get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "hESC", guides_per_junction = 2){
+  load(file = "~/isoviz/data/gencode_intron_all_data.rda")
+  gene_guides = gencode_intron_all_data %>% filter(grepl(gene, gene_id))
+  gene_name = unique(gene_guides$gene_name)
+  top_guides = gene_guides %>% filter(junction_category == "partial_unique" | junction_category == "fully_unique") %>%
+    group_by(junc_id) %>% slice_min(predicted_lfc, n = guides_per_junction)
+
+  # filter junctions based on expression or by junc_id- this is just set up to get an example
+
+  output = top_guides %>% filter(junc_id %in% for_guides$junc_id) %>% 
+    left_join(for_guides) %>%
+    select(Cluster = cluster.n, JuncID = junc_id, Category = junction_category, Counts = junc.counts, Usage = junc.usage,
+           Isoforms = isoforms, GuideID = guide_id, GuideSeq = guide_sequence, Prediction = predicted_lfc)
+
+  # output table
+  # don't add sig digits
+  fun <- function(x) {formatC(x, format = "f", digits = 0)}
+
+  return(nice_table(output, col.format.custom = 4:5, format.custom = "fun", 
+                    title = paste0(gene_name, ": Top ", guides_per_junction, " gRNAs per Junction")))
+}
+
+guide_table = get_guide_predictions("ENSG00000100320")
+
+flextable::save_as_docx(guide_table, path = "~/isoviz/plots/230420_rbfox2_guides_table.docx")
 ```
@@
 
-to_plot = map_junction("ENSG00000001167.15", cell_type = "hESC") 
+to_plot = map_junction("ENSG00000001167.15", cell_type = "hESC_med_string") 
 plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/nfya_isoforms_hESC.pdf", width = 8, height = 2)
+ggsave("~/isoviz/plots/nfya_isoforms_hESC.pdf", width = 8, height = 5)
 
@@
 
-to_plot = map_junction("ENSG00000100320", cell_type = "hESC") 
+to_plot = map_junction("ENSG00000100320", cell_type = "hESC_med_string") 
 plot_junction_to_isoform(to_plot)
@@
 
-```
+# PTBP1
+to_plot = map_junction("ENSG00000011304", cell_type = "hESC_med_string") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/ptbp1_isoforms_hESC.pdf", width = 8, height = 8.5)
 
-Generate a junction file of all possible junctions sequences for Andrew to make predictions
-- use gencode psl
-- want to make sure to have junc id that will correspond to leafcutter (can merge by sequence at first)
-- can make this information into a table
-```{r}
-load(file = "~/isoviz/data/iso_intron_data.rda")
-iso_intron_data %>% group_by(transcript_type) %>% tally()
-iso_intron_data %>% group_by(chr) %>% tally()
-iso_intron_data %>% filter(is.na(chr))
+to_plot = map_junction("ENSG00000011304", cell_type = "HEK293") 
+plot_junction_to_isoform(to_plot) # no unique junctions here, fix code
 
-# filter for protein coding and lncRNAs
-iso_intron_pc = iso_intron_data %>% filter(transcript_type == "protein_coding" | transcript_type == "lncRNA")
+to_plot = map_junction("ENSG00000011304", cell_type = "A375") 
+plot_junction_to_isoform(to_plot) # no unique junctions here
 
-# remove transcript ids and collapse on junction coords
-collapsed_junctions = iso_intron_pc %>% select(-trans_id, -transcript_name, -gene_type, -transcript_type) %>% 
-  distinct() %>% select(chr, start = intron_starts, end = intron_ends, gene_id, gene_name, strand) # 281,880 junctions- how many show any uniqueness?
+# MKNK2
+to_plot = map_junction("ENSG00000099875", cell_type = "hESC_med_string") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/mknk2_isoforms_hESC.pdf", width = 8, height = 4)
 
-# to address how many annotated isoforms have a truly unique junction
-# first filter the genes with more than one transcript
-iso_intron_pc %>% select(gene_id, trans_id) %>% distinct() # 85,586 transcripts
-n_df = iso_intron_pc %>% select(gene_id, trans_id) %>% distinct() %>% group_by(gene_id) %>% mutate(n_transcripts = n()) %>% ungroup() # 33,413
-gene_list = n_df %>% select(-trans_id) %>% distinct() %>% filter(n_transcripts > 1) %>% select(gene_id) # 17,035 genes have more than 1 isoform
+to_plot = map_junction("ENSG00000099875", cell_type = "HEK293") 
+plot_junction_to_isoform(to_plot)
 
-iso_pc_df = iso_intron_pc %>% full_join(n_df) %>% filter(gene_id %in% gene_list$gene_id) %>%
-  group_by(chr, intron_starts, intron_ends, gene_id, strand, gene_name, n_transcripts) %>% mutate(n_trans_per_junc = n(), commonality = round((n_trans_per_junc/n_transcripts)*100, 0)) %>%
-  mutate(junction_category = ifelse(n_trans_per_junc == 1, "fully_unique", ifelse(commonality == 100, "common", "partial_unique"))) %>% ungroup()
+to_plot = map_junction("ENSG00000099875", cell_type = "A375") 
+plot_junction_to_isoform(to_plot)
 
-# junction level numbers- there are ~50,000 fully unique junctions for pc genes, meaning they belong to a singular isoform
-iso_pc_df %>% select(chr, intron_starts, intron_ends, strand, gene_id, gene_name, gene_type, junction_category) %>% distinct() %>% 
-  group_by(gene_type, junction_category) %>% tally()
+# HNRNPA1
+to_plot = map_junction("ENSG00000135486", cell_type = "hESC_med_string") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/hnrnpa1_isoforms_hESC.pdf", width = 8, height = 5.5)
 
-# isoform-level, how many isoforms are uniquely targetable
-iso_pc_df %>% filter(junction_category == "fully_unique") %>% select(gene_id, gene_name, trans_id, transcript_name, gene_type) %>% distinct() # 47,657 transcripts are uniquely targetable
-iso_pc_df %>% filter(junction_category == "fully_unique") %>% select(gene_id, gene_name, trans_id, transcript_name, gene_type) %>% distinct() %>%
-  distinct(gene_name) # 16,801 genes
+# HNRNPK
+to_plot = map_junction("ENSG00000165119", cell_type = "hESC_med_string") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/hnrnpk_isoforms_hESC.pdf", width = 16, height = 6)
+
+# KHSRP- ENSG00000088247
+to_plot = map_junction("ENSG00000088247", cell_type = "hESC_med_string") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/khsrp_isoforms_hESC.pdf", width = 16, height = 6)
+
+# MATR3
+to_plot = map_junction("ENSG00000015479", cell_type = "hESC_med_string") 
+plot_junction_to_isoform(to_plot)
+ggsave("~/isoviz/plots/matr3_isoforms_hESC.pdf", width = 16, height = 6)
 

### 2023-04-24 temp_code/MS_isoviz_code.Rmd
MS test push
MS update plotting function to consider all junctions

@@
 title: "Megan's code for Isoviz"
-output: html_notebook
+
 ---
@@
 require(Biostrings)
+library(ggsci)
 require(rempsyc) # for fancy tables: https://cran.r-project.org/web/packages/rempsyc/vignettes/table.html
@@
 How do we incorporate expression here since the hESC data is more deeply sequenced- could combine feature counts when we format the leafcutter output
+Using output from minicutter
 ```{r}
-# merge with gencode info and predictions from 230411_cas13_paper_figures.Rmd
-# we should probably filter these for things like runs of T's, etc like we did when designing the screen
-# missing 166- which are guides in Harm's screen
-v41_info = read_tsv("~/isoviz/data/gencode_v41_all_junction_guides.txt", col_names = TRUE) %>% 
-  mutate(guide_sequence = as.character(guide_sequence)) # 2,255,040
-predictions = read_csv("/gpfs/commons/groups/knowles_lab/cas13_share/230414_predictions/combined-corrected-predictions-gencode.csv") %>% 
-  select(guide_id, predicted_lfc) # 2,254,874
 
-gencode_intron_all_data = v41_info %>% left_join(predictions, by = "guide_id")
+# aggregated for this function
 save(gencode_intron_all_data, file = "~/isoviz/data/gencode_intron_all_data.rda")
@@
 # map junctions function
-# I really don't need the guide info here just the junction info- consider revising and bring the guideRNAs in for another junction
 map_all_junctions = function(gene = "ENSG00000100320", cell_type = "hESC"){
@@
   # gencode- this is an updated file containing more junction info
-  load(file = "~/isoviz/data/gencode_intron_all_data.rda")
+  load(file = "~/isoviz/data/gencode_intron_all_data.rda") # 281,880
 
@@
   # this now contains all guide information for this gene
-  gene_iso_data = gencode_intron_all_data %>% filter(grepl(gene, gene_id))
+  gene_iso_data = gencode_intron_all_data %>% filter(grepl(gene, gene_id)) # 24
 
@@
   lfc_clusters = paste0(cell_type, "_all_lfc_junctions")
-  load(file = paste0("~/isoviz/data/", lfc_clusters, ".rda"))
-
-  # test
-  clusters_table %>% filter(chr == "chr22" & start > 35738735 & end < 36028824) %>% arrange(desc(start))
+  load(file = paste0("~/isoviz/data/", lfc_clusters, ".rda")) # juncs_recluster
 
   # since leafcutter junctions dont have gene annotations, need to combine here first
+  juncs_recluster %<>% select(everything(), junc.counts = readcount) %>% group_by(cluster_idx) %>% mutate(cluster.counts = sum(junc.counts)) %>% ungroup()
   gene_cluster = gene_iso_data %>% 
-    left_join(clusters_table, by = c("chr", "junc_start" = "start", "junc_end" = "end"))
+    left_join(juncs_recluster, by = c("chr" = "chrom", "junc_start" = "start", "junc_end" = "end", "strand")) %>%
+    arrange(cluster_idx) # 24
+
+  # check minicutter output
+  juncs_recluster %>% filter(cluster_idx == "134447")
 
@@
   # to get information on leafcutter junctions that don't match annotation
-  all_gene_clusters = clusters_table %>% filter(cluster.n %in% gene_cluster$cluster.n) %>% 
-    left_join(gene_iso_data, by = c("chr", "start" = "junc_start", "end" = "junc_end")) %>%
+  all_gene_clusters = juncs_recluster %>% filter(cluster_idx %in% gene_cluster$cluster_idx) %>% 
+    left_join(gene_iso_data, by = c("chrom" = "chr", "start" = "junc_start", "end" = "junc_end", "strand")) %>%
     filter(is.na(gene_id)) %>% 
@@
            gene_id = unique(gene_cluster$gene_id), junction_category = "unknown", 
-           transcript_isoforms = "unknown", junc_id = paste0("unk", 1:nrow(.))) %>%
-    select(chr, junc_start = start, junc_end = end, everything()) %>%
+           transcript_isoforms = "unknown", junc_id = paste0("unk.", row_number())) %>%
+    select(chr = chrom, junc_start = start, junc_end = end, everything()) %>%
     bind_rows(gene_cluster) %>%
-    mutate(junc.counts = ifelse(is.na(cluster.n), 0, junc.counts), cluster.counts = ifelse(is.na(cluster.n), 0, cluster.counts))
+    mutate(junc.counts = ifelse(is.na(cluster_idx), 0, junc.counts), cluster.counts = ifelse(is.na(cluster_idx), 0, cluster.counts)) %>%
+    arrange(cluster_idx)
 
-  n = all_gene_clusters %>% distinct(cluster.n, junc_id) %>% group_by(cluster.n) %>% 
-    mutate(junc.per.cluster = ifelse(is.na(cluster.n), 1, n())) %>% ungroup()
+  n = all_gene_clusters %>% distinct(cluster_idx, junc_id) %>% group_by(cluster_idx) %>% 
+    mutate(junc.per.cluster = ifelse(is.na(cluster_idx), 1, n())) %>% ungroup()
 
-  all_gene_clusters %<>% left_join(n, by = c("cluster.n", "junc_id")) %>%
+  all_gene_clusters %<>% left_join(n, by = c("cluster_idx", "junc_id")) %>%
     mutate(junc.usage = ifelse(junc.counts == 0, 0, round((junc.counts/cluster.counts)*100, digits = 0)), cell_line = paste0(cell_type))
@@
   #test
-  all_gene_clusters %>% filter(grepl("209", transcript_isoforms)) %>%
-    select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
-    distinct() %>% arrange(junc_start)
+  all_gene_clusters %>% filter(grepl("209", transcript_isoforms)) %>% arrange(junc_start)
 
@@
   # I want to be careful here since some junctions won't be represented because of technical issues so only eliminate isoforms where none of their unique junctions have counts, avoid filtering on partially unique for now
-  non_expressed_genes = all_gene_clusters %>% select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
-    distinct() %>% separate_longer_delim(transcript_isoforms, delim = ",") %>% group_by(transcript_isoforms, junction_category) %>%
+  non_expressed_genes = all_gene_clusters %>%
+    separate_longer_delim(transcript_isoforms, delim = ",") %>% group_by(transcript_isoforms, junction_category) %>%
     mutate(isoform_counts = ifelse(junction_category == "fully_unique", sum(junc.counts), junc.counts)) %>% 
@@
 
-  uniquely_targetable = all_gene_clusters %>% select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
-    distinct() %>% separate_longer_delim(transcript_isoforms, delim = ",") %>% filter(junction_category == "fully_unique") %>%
+  uniquely_targetable = all_gene_clusters %>% 
+    separate_longer_delim(transcript_isoforms, delim = ",") %>% filter(junction_category == "fully_unique") %>%
     distinct(transcript_isoforms)
 
-  partially_targetable = all_gene_clusters %>% select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
-    distinct() %>% separate_longer_delim(transcript_isoforms, delim = ",") %>% filter(junction_category == "partial_unique") %>%
+  partially_targetable = all_gene_clusters %>% 
+    separate_longer_delim(transcript_isoforms, delim = ",") %>% filter(junction_category == "partial_unique") %>%
     distinct(transcript_isoforms)
 
-  output = all_gene_clusters %>% select(-guide_start, -guide_end, -guide_sequence, -target_sequence, -guide_id, -predicted_lfc) %>%
-    distinct() %>% separate_longer_delim(transcript_isoforms, delim = ",") %>%
+  output = all_gene_clusters %>% 
+    separate_longer_delim(transcript_isoforms, delim = ",") %>%
     mutate(transcript_targetable = ifelse(transcript_isoforms %in% uniquely_targetable$transcript_isoforms, "Uniquely", 
                                           ifelse(transcript_isoforms %in% partially_targetable$transcript_isoforms, "Partially", "None"))) %>% 
-    mutate(isoform_expressed = ifelse(transcript_isoforms %in% non_expressed_genes$transcript_isoforms, "Unlikely", "Likely"))
+    mutate(isoform_expressed = ifelse(transcript_isoforms %in% non_expressed_genes$transcript_isoforms, "Unlikely", "Likely")) # 135
 
@@
 
+# Outdated
 ```{r}
 load(file='~/isoviz/data/iso_data_complete.rda')
-#df = to_plot
 
@@
 Updated plotting to go with all junctions output
+Now includes minicutter output so all junctions represented
 ```{r}
-# might have to use regtools if I truly want all junctions
 load(file='~/isoviz/data/iso_data_complete.rda')
-df = test
+#df = test
 
 # plot isoform plus leafcutter clusters
-plot_all_junction_to_isoform = function(df, junction_usage = 5, include_common_juncs = FALSE, filter_expressed_isoforms = TRUE){
+# add a parameter to only plot a specific junction
+plot_all_junction_to_isoform = function(gene = "ENSG00000001167.15", cell_type = "hESC", junction_usage = 5, include_common_juncs = FALSE, filter_expressed_isoforms = TRUE){
 
-  # filter here based on input parameters
+  # call map_junction function here
+  df = map_all_junctions(gene, cell_type)
+
+  # filter here based on input parameters- also unknown junctions that belong to common junctions
   if(include_common_juncs == FALSE){
-    df %<>% filter(junction_category != "common")
+    remove_cluster = df %>% filter(junction_category == "common") %>% select(cluster_idx) %>% distinct()
+    df %<>% filter(!cluster_idx %in% remove_cluster$cluster_idx)
   }
@@
     df %<>% filter(isoform_expressed == "Likely")
+
+     # create list of expressed transcripts
+    expressed_isoforms = df %>% filter(isoform_expressed == "Likely", transcript_isoforms != "unknown") %>% 
+      separate(col = transcript_isoforms, into = c("name", "transcript")) %>% select(transcript) %>% distinct() %>%
+      arrange(transcript) %>%
+      mutate(isoforms = paste0(transcript, collapse = ",")) %>% select(isoforms) %>% distinct()
   }
@@
     xlim(min_start, max_end) + ylim(0,y_max) +
-    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$transcript_name), size = 3) +
+    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = -0.1, label = to_plot_ordered$transcript_name), size = 3, check_overlap = TRUE) +
     theme_bw() + ylab("") + xlab("") +
@@
           plot.title = element_text(hjust = 0.5), plot.margin = margin(0.1,1,0,0.1, "in")) +
-    ggtitle(paste0(name, " Junction to Isoform Map (" , to_plot_ordered$strand[1], " strand )")) + 
+    ggtitle(paste0(cell_type, ": ", name, " Junction to Isoform Map (" , to_plot_ordered$strand[1], " strand )")) + 
     coord_cartesian(xlim = c(min_start, max_end), clip = 'off')
@@
     mutate(transcript = ifelse(is.na(transcript), "unknown", transcript)) %>%
-    group_by(cluster.n, junc_id) %>% 
-    arrange(cluster.n, junc_id, transcript) %>%
+    group_by(cluster_idx, junc_id) %>% 
+    arrange(cluster_idx, junc_id, transcript) %>%
     mutate(isoforms = paste0(transcript, collapse = ",")) %>%
     select(1:5, junc_id, junction_category, junc.usage, isoforms, junc.per.cluster) %>% distinct() %>% ungroup() %>% 
-    mutate(text_plot = paste0(junc.counts, " reads (", junc.usage, "%) : ", isoforms))
+    mutate(text_plot = paste0(junc.counts, " reads (", junc.usage, "%) : ", isoforms), alpha = junc.usage/100)
 
-  clusters = unique(df$cluster.n)
+  clusters = unique(df$cluster_idx)
   introns = df %>% select(-gene_id, -gene_name, -transcript_isoforms) %>% 
-    distinct() %>% arrange(cluster.n, junc_start)
+    distinct() %>% arrange(cluster_idx, junc.usage)
+  introns %<>% left_join(text)
+
+  # gets rid of gencode partially unique junctions that are common in this cell type (100% usage) 
+  if(filter_expressed_isoforms == TRUE){
+    remove_junc = introns %>% filter(junc.usage & isoforms %in% expressed_isoforms$isoforms)
+    introns %<>% filter(!junc_id %in% remove_junc$junc_id)
+  }
+
   introns$junc.order = 1:nrow(introns)
-  introns %<>% left_join(text)
 
@@
   p2 = ggplot() + 
-    geom_rect(data = introns, mapping = aes(xmin = junc_start, xmax = junc_end, ymin = junc.order - 0.2, ymax = junc.order + 0.2, fill = cluster.n)) + 
-    geom_text(data = introns, aes(x = junc_end, y = junc.order, label = text_plot), hjust = 0, size = 3) +
+    geom_rect(data = introns, mapping = aes(xmin = junc_start, xmax = junc_end, ymin = junc.order - 0.2, ymax = junc.order + 0.2, fill = as.character(cluster_idx), alpha = alpha)) + 
+    geom_text(data = introns, aes(x = junc_end, y = junc.order, label = text_plot), hjust = "inward", size = 3) +
+    scale_fill_npg() + 
     xlim(min_start, max_end) + theme_bw() + 
     xlab("Hg38 Genomic Position (bp)") + ylab("") + 
-    geom_text(data = introns, aes(x = Inf, y = junc.order, hjust = 0, label = junc_id), size = 3) + 
+    geom_text(data = introns, aes(x = Inf, y = junc.order, hjust = -0.1, label = junc_id), size = 3) + 
     theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "none", plot.margin = margin(0,1,0,0.1, "in")) +
@@
   t = (n + nrow(introns))
-  h = t/2
-  p1_p2 = plot_grid(p1, p2, ncol = 1, align = "v", rel_heights = c(n*0.8, nrow(introns)))
+  h = t/3
+
+  # p1
+  if(n <= 4){
+    iso_h = n*1.3
+  }else{
+    iso_h = n*0.8
+  }
+
+  # p2
+  if(nrow(introns) > 6){
+    junc_h = nrow(introns) * 0.8
+  }else{
+    junc_h = nrow(introns) * 1.2
+  }
+  p1_p2 = plot_grid(p1, p2, ncol = 1, align = "v", rel_heights = c(iso_h, junc_h))
 
@@
 
-ggsave("~/isoviz/plots/rbfox2_test.pdf", width = 8, height = h)
+#map_all_junctions("ENSG00000001167.15", cell_type = "hESC")
+plot_all_junction_to_isoform("ENSG00000100320", cell_type = "hESC") # RBFOX2
+ggsave("~/isoviz/plots/rbfox2_hESC_filtered.png", width = 8, height = 5)
+
+plot_all_junction_to_isoform("ENSG00000001167.15", cell_type = "HEK293") # NFYA
+ggsave("~/isoviz/plots/nfya_hek293_filtered.png", width = 8, height = 3)
 
@@
 ```{r}
+# merge with gencode info and predictions from 230411_cas13_paper_figures.Rmd
+# we should probably filter these for things like runs of T's, etc like we did when designing the screen
+# missing 166- which are guides in Harm's screen
+v41_info = read_tsv("~/isoviz/data/gencode_v41_all_junction_guides.txt", col_names = TRUE) %>% 
+  mutate(guide_sequence = as.character(guide_sequence)) # 2,255,040
+predictions = read_csv("/gpfs/commons/groups/knowles_lab/cas13_share/230414_predictions/combined-corrected-predictions-gencode.csv") %>% 
+  select(guide_id, predicted_lfc) # 2,254,874
+
+gencode_predictions = v41_info %>% left_join(predictions, by = "guide_id")
+#save(gencode_predictions, file = "~/isoviz/data/gencode_predictions.rda")
+
 # need to read in count information a different way- this is just a quick and dirty method
 # write a function to get expressed junctions/isoforms and call that within functions?
-get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "hESC", guides_per_junction = 2){
-  load(file = "~/isoviz/data/gencode_intron_all_data.rda")
-  gene_guides = gencode_intron_all_data %>% filter(grepl(gene, gene_id))
-  gene_name = unique(gene_guides$gene_name)
-  top_guides = gene_guides %>% filter(junction_category == "partial_unique" | junction_category == "fully_unique") %>%
-    group_by(junc_id) %>% slice_min(predicted_lfc, n = guides_per_junction)
+# some of this is copied from map_all_junctions so should make this a function
+junctions = c("junc178149", "junc178148")
+junc_list = junctions
+get_guide_predictions = function(junc_list, cell_type = "hESC", guides_per_junction = 2){
+  load(file = "~/isoviz/data/gencode_predictions.rda")
+  filtered_guides = gencode_predictions %>% filter(junc_id %in% junc_list)
+  gene_name = unique(filtered_guides$gene_name)
+  gene = unique(filtered_guides$gene_id)
 
-  # filter junctions based on expression or by junc_id- this is just set up to get an example
+  # start function here
+  load(file = "~/isoviz/data/gencode_intron_all_data.rda") # 281,880
+  gene_iso_data = gencode_intron_all_data %>% filter(grepl(gene, gene_id))
 
-  output = top_guides %>% filter(junc_id %in% for_guides$junc_id) %>% 
-    left_join(for_guides) %>%
-    select(Cluster = cluster.n, JuncID = junc_id, Category = junction_category, Counts = junc.counts, Usage = junc.usage,
-           Isoforms = isoforms, GuideID = guide_id, GuideSeq = guide_sequence, Prediction = predicted_lfc)
+  lfc_clusters = paste0(cell_type, "_all_lfc_junctions")
+  load(file = paste0("~/isoviz/data/", lfc_clusters, ".rda")) # juncs_recluster
+
+  # since leafcutter junctions dont have gene annotations, need to combine here first
+  juncs_recluster %<>% select(everything(), junc.counts = readcount) %>% group_by(cluster_idx) %>% mutate(cluster.counts = sum(junc.counts)) %>% ungroup()
+  gene_cluster = gene_iso_data %>% 
+    left_join(juncs_recluster, by = c("chr" = "chrom", "junc_start" = "start", "junc_end" = "end", "strand")) %>%
+    arrange(cluster_idx) # 24
+
+  # to get information on leafcutter junctions that don't match annotation
+  all_gene_clusters = juncs_recluster %>% filter(cluster_idx %in% gene_cluster$cluster_idx) %>% 
+    left_join(gene_iso_data, by = c("chrom" = "chr", "start" = "junc_start", "end" = "junc_end", "strand")) %>%
+    filter(is.na(gene_id)) %>% 
+    mutate(gene_name = unique(gene_cluster$gene_name), strand = unique(gene_cluster$strand), 
+           gene_id = unique(gene_cluster$gene_id), junction_category = "unknown", 
+           transcript_isoforms = "unknown", junc_id = paste0("unk.", row_number())) %>%
+    select(chr = chrom, junc_start = start, junc_end = end, everything()) %>%
+    bind_rows(gene_cluster) %>%
+    mutate(junc.counts = ifelse(is.na(cluster_idx), 0, junc.counts), cluster.counts = ifelse(is.na(cluster_idx), 0, cluster.counts)) %>%
+    arrange(cluster_idx)
+
+  n = all_gene_clusters %>% distinct(cluster_idx, junc_id) %>% group_by(cluster_idx) %>% 
+    mutate(junc.per.cluster = ifelse(is.na(cluster_idx), 1, n())) %>% ungroup()
+
+  all_gene_clusters %<>% left_join(n, by = c("cluster_idx", "junc_id")) %>%
+    mutate(junc.usage = ifelse(junc.counts == 0, 0, round((junc.counts/cluster.counts)*100, digits = 0)), cell_line = paste0(cell_type))
+  # end here
+
+  top_guides = filtered_guides %>%
+    group_by(junc_id) %>% slice_min(predicted_lfc, n = guides_per_junction) %>% ungroup()
+
+  # make the isoforms shorthand- probably should change in original file
+  output = top_guides %>% left_join(all_gene_clusters) %>%
+    select(Cluster = cluster_idx, JuncID = junc_id, Category = junction_category, Counts = junc.counts, Usage = junc.usage,
+           Isoforms = transcript_isoforms, GuideID = guide_id, GuideSeq = guide_sequence, Prediction = predicted_lfc)
 
@@
 
-  return(nice_table(output, col.format.custom = 4:5, format.custom = "fun", 
+  return(nice_table(output, col.format.custom = 1:5, format.custom = "fun", 
                     title = paste0(gene_name, ": Top ", guides_per_junction, " gRNAs per Junction")))

### 2023-04-28 temp_code/MS_isoviz_code.Rmd
MS update grna prediction

@@
 Inputs leafcutter clustering output and outputs formatted table
+This is old since we use minicutter now
 ```{r}
@@
 # Consider all junction counts and use gencode annotations to define unique versus common
+# this is also irrelevant now since we have minicutter
 ```{r}
@@
 ```
+
+### Potential function #1: ###
 
@@
 
-Function: Map junction to isoform
-```{r}
-load(file = "~/isoviz/data/iso_intron_data.rda")
-#load(file = "~/isoviz/data/a375_screen_lfc_clusters.rda")
-#load(file = "~/isoviz/data/HEK293_lfc_clusters.rda")
-
-# is it better to load rda within a function?
-# map junctions function- for now, it isn't using cell type but eventually we should change this
-map_junction = function(gene = "ENSG00000001167.15", cell_type = "HEK293"){
-
-  # filter for gene, for now, also filter for protein coding transcripts- can set this as a parameter later
-  gene_iso_data = iso_intron_data %>% filter(grepl(gene, gene_id), transcript_type == "protein_coding") # 17 for NFYA
-
-  # join iso data with leafcutter data and mark any unannotated junctions
-  # consider cell type
-  lfc_clusters = paste0(cell_type, "_lfc_clusters")
-  load(file = paste0("~/isoviz/data/", lfc_clusters, ".rda"))
-  const_junc_list = clusters_table %>% 
-    group_by(cluster.n) %>% 
-    tally() %>% filter(n == 1) %>% ungroup()
-  clusters_table %<>% mutate(junc.type = ifelse(cluster.n %in% const_junc_list$cluster.n, "Common", "Unique"))
-  gene_cluster = gene_iso_data %>% 
-    left_join(clusters_table, by = c("chr", "intron_starts" = "start", "intron_ends" = "end")) %>%
-    filter(!is.na(cluster.n))
-
-  gene = unique(gene_cluster$gene_id)
-  name = unique(gene_cluster$gene_name)
-  gene_strand = unique(gene_cluster$strand)
-  cluster_ids = unique(gene_cluster$cluster.n)
-
-  output = clusters_table %>% filter(cluster.n %in% cluster_ids) %>% group_by(cluster.n) %>% mutate(junc.per.cluster = n()) %>% 
-    left_join(gene_iso_data, by = c("chr", "start" = "intron_starts", "end" = "intron_ends")) %>% 
-    mutate(junc.usage = round((junc.counts/cluster.counts)*100, digits = 0), 
-           transcript_name = ifelse(is.na(transcript_name), paste0(cluster.n, "_Unannotated"), transcript_name),
-           trans_id = ifelse(is.na(trans_id), paste0(cluster.n, "_Unannotated"), trans_id),
-           gene_id = ifelse(is.na(gene_id), gene, gene_id),
-           gene_name = ifelse(is.na(gene_name), name, gene_name),
-           strand = ifelse(is.na(strand), gene_strand, strand), cell_line = paste0(cell_type))
-
-  return(output)
-}
-
-#to_plot = map_junction("ENSG00000001167.15") # NFYA
-#to_plot = map_junction("ENSG00000100320") # RBFOX2, alt start is missing because it's intron doesnt overlap another and counts are only 8 so included in consitutive either
-
-# can we ask if all junctions of an isoform are represented to filter isoforms?
-# can we create pseudo regtools output based on all possible junctions and run that through leafcutter to get true common versus unique
-
-```
-
 Updated Function: Map junction to isoform: Updated to include all junctions and use gencode annotations to mark unique versus common
@@
 
-# Outdated
-```{r}
-load(file='~/isoviz/data/iso_data_complete.rda')
-
-# plot isoform plus leafcutter clusters
-plot_junction_to_isoform = function(df, include_common_juncs = FALSE){
-
-  # I am copying the plot_isoforms function here, figure out how to integrate these later
-  gene = unique(df$gene_id)
-  gene_trans = which(str_detect(iso_data_complete$gene_id, gene))
-  gene_data = iso_data_complete[gene_trans,]
-
-  # filter out common junctions for plotting (default)
-  if(include_common_juncs == FALSE){
-    df %<>% filter(junc.type == "Unique")
-  }
-
-  # get info
-  name = unique(df$gene_name)
-  transcripts = unique(df$transcript_name)
-
-  # filter for transcripts that have junction level data to support
-  gene_data %<>% filter(transcript_name %in% transcripts)
-
-  # setting parameters for the plot based on length of the gene and number of isoforms
-  length = max(gene_data$end) - min(gene_data$start)
-  min_start = min(gene_data$start)
-  n = gene_data %>% group_by(trans_id) %>% tally() %>% nrow()
-  y_max = n + 1
-  max_end = max(gene_data$end)
-
-  # order the isoforms by length for plotting
-  order_plot = gene_data %>% group_by(transcript_name, transcript_length) %>% tally() %>%
-    dplyr::arrange(desc(transcript_length), desc(n))
-  order_plot$trans_order = 1:nrow(order_plot)
-  order_plot = order_plot %>% dplyr::select(transcript_name, trans_order)
-
-  to_plot_ordered = gene_data %>% full_join(order_plot, by = c("transcript_name")) %>%
-    arrange(trans_order, blockstarts) 
-  to_plot_ordered = to_plot_ordered %>% mutate(seg_end = end)
-
-  # make isoform plot
-  p1 = ggplot() +
-    geom_segment(aes(x = to_plot_ordered$start, y = to_plot_ordered$trans_order,
-                     xend = to_plot_ordered$seg_end, yend = to_plot_ordered$trans_order)) +
-    geom_rect(data = to_plot_ordered, mapping = aes(xmin = blockstarts, xmax = blockends, ymin = trans_order - 0.3, ymax = trans_order + 0.3)) +
-    xlim(min_start, max_end) + ylim(0,y_max) +
-    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$transcript_name), size = 3) +
-    theme_bw() + ylab("") + xlab("") +
-    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), 
-          axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "top", 
-          plot.title = element_text(hjust = 0.5), plot.margin = margin(0.1,1,0,0.1, "in")) +
-    ggtitle(paste0(name, " Junction to Isoform Map (" , to_plot_ordered$strand[1], " strand )")) + 
-    coord_cartesian(xlim = c(min_start, max_end), clip = 'off')
-
-  # new code to plot leafcutter cluster info underneath isoform level data
-  # aggregate isoform information
-  text = df %>% separate(col = transcript_name, into = c("name", "transcript")) %>% group_by(cluster.n, start, end) %>% 
-    arrange(cluster.n, start, end, transcript) %>%
-    mutate(isoforms = paste0(transcript, collapse = ",")) %>%
-    select(1:5, junc.usage, isoforms) %>% distinct() %>% ungroup() %>% 
-    mutate(text_plot = paste0(junc.counts, " reads (", junc.usage, "%) : ", isoforms))
-
-  clusters = unique(df$cluster.n)
-  introns = df %>% select(-gene_id, -trans_id, -gene_name, -transcript_name) %>% 
-    distinct() %>% arrange(cluster.n, start)
-  introns$junc.order = 1:nrow(introns)
-  introns %<>% left_join(text)
-  p2 = ggplot() + 
-    geom_rect(data = introns, mapping = aes(xmin = start, xmax = end, ymin = junc.order - 0.2, ymax = junc.order + 0.2, fill = cluster.n)) + 
-    geom_text(data = introns, aes(x = end, y = junc.order, label = text_plot), hjust = 0, size = 3) +
-    xlim(min_start, max_end) + theme_bw() + 
-    xlab("Hg38 Genomic Position (bp)") + ylab("") + 
-    theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "none", plot.margin = margin(0,1,0,0.1, "in")) +
-    coord_cartesian(xlim = c(min_start, max_end))
-
-  # plot dimenstions based on number of isoforms and junctions
-  t = (n + nrow(introns))
-  h = t/2
-  p1_p2 = plot_grid(p1, p2, ncol = 1, align = "v", rel_heights = c(n*0.8, nrow(introns)))
-  #ggsave("~/isoviz/plots/rbfox2_isoforms.pdf", width = 8, height = h)
-  return(p1_p2)
-}
-
-#plot_junction_to_isoform(to_plot)
-#ggsave("~/isoviz/plots/rbfox2_isoforms.pdf", width = 8, height = 8.5)
-```
-
 Updated plotting to go with all junctions output
@@
 # add a parameter to only plot a specific junction
-plot_all_junction_to_isoform = function(gene = "ENSG00000001167.15", cell_type = "hESC", junction_usage = 5, include_common_juncs = FALSE, filter_expressed_isoforms = TRUE){
+plot_all_junction_to_isoform = function(gene = "ENSG00000099875", cell_type = "hESC", junction_usage = 5, include_common_juncs = FALSE, filter_expressed_isoforms = TRUE, filter_all_100percent = FALSE){
 
@@
   # gets rid of gencode partially unique junctions that are common in this cell type (100% usage) 
-  if(filter_expressed_isoforms == TRUE){
-    remove_junc = introns %>% filter(junc.usage & isoforms %in% expressed_isoforms$isoforms)
+  if(filter_expressed_isoforms == TRUE & filter_all_100percent == FALSE){
+    remove_junc = introns %>% filter(junc.usage == "100" & isoforms %in% expressed_isoforms$isoforms)
     introns %<>% filter(!junc_id %in% remove_junc$junc_id)
+  }else{
+    introns %<>% filter(junc.usage != "100")
   }
@@
 
+plot_all_junction_to_isoform("ENSG00000099875", cell_type = "hESC", filter_all_100percent = TRUE) # MKNK2
+ggsave("~/isoviz/plots/mknk2_filtered.png", width = 8, height = 3)
+
 #df %>% filter(transcript_isoforms == "RBFOX2-209") %>% arrange(junc_start)
@@
 junctions = c("junc178149", "junc178148")
+junctions = c("junc128207", "junc128208")
+
 junc_list = junctions
@@
     select(Cluster = cluster_idx, JuncID = junc_id, Category = junction_category, Counts = junc.counts, Usage = junc.usage,
-           Isoforms = transcript_isoforms, GuideID = guide_id, GuideSeq = guide_sequence, Prediction = predicted_lfc)
-
+          GuideSeq = guide_sequence, Prediction = predicted_lfc)
+  #select(Cluster = cluster_idx, JuncID = junc_id, Category = junction_category, Counts = junc.counts, Usage = junc.usage,Isoforms = transcript_isoforms, GuideID = guide_id, GuideSeq = guide_sequence, Prediction = predicted_lfc)
   # output table
@@
 ```
-
-```{r}
-# NFYA
-to_plot = map_junction("ENSG00000001167.15", cell_type = "A375") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/nfya_isoforms.pdf", width = 8, height = 4)
-
-to_plot = map_junction("ENSG00000001167.15", cell_type = "HEK293") 
-plot_junction_to_isoform(to_plot)
-# for height, count rows and divide by 2
-ggsave("~/isoviz/plots/nfya_isoforms_hek293.pdf", width = 8, height = 2)
-
-to_plot = map_junction("ENSG00000001167.15", cell_type = "hESC_med_string") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/nfya_isoforms_hESC.pdf", width = 8, height = 5)
-
-# RBFOX2
-to_plot = map_junction("ENSG00000100320", cell_type = "A375") # alt start is missing because it's intron doesnt overlap another and counts are only 8 so included in consitutive either
-plot_junction_to_isoform(to_plot)
-
-to_plot = map_junction("ENSG00000100320", cell_type = "HEK293") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/rbfox2_isoforms_hek293.pdf", width = 8, height = 9)
-
-to_plot = map_junction("ENSG00000100320", cell_type = "hESC_med_string") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/rbfox2_isoforms_hESC.pdf", width = 8, height = 8.5)
-
-# PTBP1
-to_plot = map_junction("ENSG00000011304", cell_type = "hESC_med_string") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/ptbp1_isoforms_hESC.pdf", width = 8, height = 8.5)
-
-to_plot = map_junction("ENSG00000011304", cell_type = "HEK293") 
-plot_junction_to_isoform(to_plot) # no unique junctions here, fix code
-
-to_plot = map_junction("ENSG00000011304", cell_type = "A375") 
-plot_junction_to_isoform(to_plot) # no unique junctions here
-
-# MKNK2
-to_plot = map_junction("ENSG00000099875", cell_type = "hESC_med_string") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/mknk2_isoforms_hESC.pdf", width = 8, height = 4)
-
-to_plot = map_junction("ENSG00000099875", cell_type = "HEK293") 
-plot_junction_to_isoform(to_plot)
-
-to_plot = map_junction("ENSG00000099875", cell_type = "A375") 
-plot_junction_to_isoform(to_plot)
-
-# HNRNPA1
-to_plot = map_junction("ENSG00000135486", cell_type = "hESC_med_string") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/hnrnpa1_isoforms_hESC.pdf", width = 8, height = 5.5)
-
-# HNRNPK
-to_plot = map_junction("ENSG00000165119", cell_type = "hESC_med_string") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/hnrnpk_isoforms_hESC.pdf", width = 16, height = 6)
-
-# KHSRP- ENSG00000088247
-to_plot = map_junction("ENSG00000088247", cell_type = "hESC_med_string") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/khsrp_isoforms_hESC.pdf", width = 16, height = 6)
-
-# MATR3
-to_plot = map_junction("ENSG00000015479", cell_type = "hESC_med_string") 
-plot_junction_to_isoform(to_plot)
-ggsave("~/isoviz/plots/matr3_isoforms_hESC.pdf", width = 16, height = 6)
-
-```

### 2023-04-28 temp_code/minicutter.R

@@ -1,0 +1,81 @@
+# pulled from David's github on 4/20/23
+# edited regtools input on 4/21/23
+
+require(tidyverse)
+require(Matrix)
+require(igraph)
+
+leafcutter_one_step = function(juncs) {
+
+  splice_sites = bind_rows(
+    juncs %>% select(chrom, strand, position = start),
+    juncs %>% select(chrom, strand, position = end)) %>%
+    distinct() %>%
+    arrange(chrom, strand, position) %>%
+    mutate(idx = 1:n())
+
+  juncs = juncs %>%
+    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", start = "position")) %>%
+    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", end = "position"), suffix = c("_start","_end"))
+
+  nss = nrow(splice_sites)
+
+  intron_connectivity <- sparseMatrix(
+    i = juncs$idx_start,
+    j = juncs$idx_end,
+    dims = c(nss, nss),
+    x = 1,
+    symmetric = TRUE
+  )
+
+  g = graph_from_adjacency_matrix(intron_connectivity, "undirected")
+
+  components(g)$membership[juncs$idx_start]
+}
+
+leafcutter_refine = function(juncs, min_usage_ratio = 0.001) {
+  juncs_filtered = juncs %>% filter(usage_ratio >= min_usage_ratio)
+  juncs_filtered$cluster_idx = leafcutter_one_step(juncs_filtered)
+  juncs_filtered
+}
+
+calculate_usage_ratios = function(juncs) {
+  juncs %>% group_by(cluster_idx) %>% mutate(usage_ratio = readcount / sum(readcount)) %>% ungroup() # calculate usage ratios
+}
+
+leafcutter_two_step = function(juncs, min_usage_ratio = 0.001) {
+  juncs$cluster_idx = leafcutter_one_step(juncs) # get inital clustering
+  juncs = juncs %>% calculate_usage_ratios()
+  leafcutter_refine(juncs, min_usage_ratio = min_usage_ratio)
+}
+
+plot_cluster_sizes = function(juncs) {
+  cluster_sizes = juncs %>% group_by(cluster_idx) %>% summarize(n = n()) %>% ungroup()
+  ta = table(cluster_sizes$n)
+  tibble(
+    cluster_size = as.numeric(names(ta)),
+    num_clusters = as.numeric(ta)
+  ) %>%
[…]

### 2023-05-07 R/isoviz_coords.R
uploaded function and start vignette

@@ -1,0 +1,85 @@
+#' A function for obtaining all necessary coordinates for leafcutter integration
+#' and plotting isoforms given a gtf file or psl file from long read sequencing
+#'
+#' Inputs gencode gtf or psl file, outputs exon coords (for plotting) and
+#' intron coords (for leafcutter integration) with ENSG/ENST names plus
+#' meaningful names (ex RBFOX2)
+#' @param file_name Gencode gtf or psl file
+#' @return list of exon and intron coordinates in that order.
+#' @examples
+#' isoviz_coords("data-raw/gencode.v41.basic.annotation.psl")
+#' @name isoviz_coords
+
+isoviz_coords = function(file_path,
+                         input_type="psl"){
+
+  if(input_type=="psl"){
+
+    #print(paste("The input file being used is"), file_path)
+    genome_data <- fread(file_path)
+    genome_data$gene_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][2]})
+    genome_data$trans_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][1]})
+
+    # Get coordinates for each "block"
+    genome_data$blocksizes = sapply(genome_data$V19, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
+    genome_data$blockstarts = sapply(genome_data$V21, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
+    genome_data$strand = genome_data$V9
+    genome_data$chr = genome_data$V14
+    genome_data$start = genome_data$V16
+    genome_data$end = genome_data$V17
+    genome_data$transcript_length = genome_data$end - genome_data$start
+
+    # Clean up and seperate blocks into rows
+    # Need to do different adjustments here to get intron coords that will line up with leafcutter
+    genome_data = genome_data %>% dplyr::select(chr, start, end, trans_id, gene_id,strand, blocksizes, blockstarts,
+                                            transcript_length) %>% separate_rows(blocksizes, blockstarts)
+    genome_data$blockstarts=as.numeric(genome_data$blockstarts)
+    genome_data$blocksizes=as.numeric(genome_data$blocksizes)
+    genome_data$blockends = genome_data$blockstarts + genome_data$blocksizes
+
+    print(paste(length(unique(genome_data$trans_id)), "transcripts with at least one intron!"))
+
+    # for each transcript
[…]

### 2023-05-07 R/plot_isoforms.R

@@ -11,7 +11,7 @@ load(file='data/iso_exon_data.rda')
 
 plot_isoforms = function(gene="ENSG00000100320"){
 
-  gene_trans =which(str_detect(genomedata$gene_id, gene))
+  gene_trans = which(str_detect(genomedata$gene_id, gene))
   gene_data = genomedata[gene_trans,]
 
   # setting parameters for the plot based on length of the gene and number of isoforms

### 2023-05-08 R/isoviz_coords.R
uploaded function and start vignette

@@ -16,7 +16,7 @@ isoviz_coords = function(file_path,
   if(input_type=="psl"){
 
     #print(paste("The input file being used is"), file_path)
-    genome_data <- fread(file_path)
+    genome_data <- data.table::fread(file_path)
     genome_data$gene_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][2]})
     genome_data$trans_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][1]})
 

### 2023-05-08 R/isoviz_coords.R
uploaded function and start vignette

@@ -9,14 +9,15 @@
 #' @examples
 #' isoviz_coords("data-raw/gencode.v41.basic.annotation.psl")
 #' @name isoviz_coords
+#' @import data.table
 
-isoviz_coords = function(file_path,
+isoviz_coords = function(file_path, gene_trans,
                          input_type="psl"){
 
   if(input_type=="psl"){
 
     #print(paste("The input file being used is"), file_path)
-    genome_data <- data.table::fread(file_path)
+    genome_data <- fread(file_path)
     genome_data$gene_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][2]})
     genome_data$trans_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][1]})
 
@@ -63,7 +64,7 @@ isoviz_coords = function(file_path,
     print(paste(length(unique(intron_data$trans_id)), "transcripts with at least one intron!"))
 
     # Convert to gene and transcript names (should this be placed somewhere else?)
-    convert = read_tsv("data-raw/gencode_v41_gene-transcript-convert.txt", col_names = TRUE)
+    convert = read_tsv(gene_trans, col_names = TRUE)
 
     # Should avoid selecting columns this way
     trans_info = genome_data %>% select(1, 4, 5, 6) %>% distinct() %>% filter(chr != "chrY", chr != "chrM")

### 2023-05-08 R/isoviz_coords.R
fixing exports

@@ -4,12 +4,14 @@
 #' Inputs gencode gtf or psl file, outputs exon coords (for plotting) and
 #' intron coords (for leafcutter integration) with ENSG/ENST names plus
 #' meaningful names (ex RBFOX2)
-#' @param file_name Gencode gtf or psl file
+#' @param file_path Gencode gtf or psl file
+#' @param gene_trans Dataframe with ensembl gene and transcript ids and names
 #' @return list of exon and intron coordinates in that order.
 #' @examples
-#' isoviz_coords("data-raw/gencode.v41.basic.annotation.psl")
+#' isoviz_coords()
 #' @name isoviz_coords
 #' @import data.table
+#' @export
 
 isoviz_coords = function(file_path, gene_trans,
                          input_type="psl"){

### 2023-05-08 R/minicutter.R

@@ -1,0 +1,40 @@
+#' A function for obtaining leafcutter intron clusters
+#'
+#' Inputs junction coordinates from regtools
+#' @param juncs Gencode gtf or psl file
+#' @return intron clusters
+#' @examples
+#' minicutter()
+#' @name minicutter
+#' @import tidyverse
+#' @import Matrix
+#' @import igraph
+
+leafcutter_one_step = function(juncs) {
+
+  splice_sites = bind_rows(
+    juncs %>% select(chrom, strand, position = start),
+    juncs %>% select(chrom, strand, position = end)) %>%
+    distinct() %>%
+    arrange(chrom, strand, position) %>%
+    mutate(idx = 1:n())
+
+  juncs = juncs %>%
+    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", start = "position")) %>%
+    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", end = "position"), suffix = c("_start","_end"))
+
+  nss = nrow(splice_sites)
+
+  intron_connectivity <- sparseMatrix(
+    i = juncs$idx_start,
+    j = juncs$idx_end,
+    dims = c(nss, nss),
+    x = 1,
+    symmetric = TRUE
+  )
+
+  g = graph_from_adjacency_matrix(intron_connectivity, "undirected")
+
+  components(g)$membership[juncs$idx_start]
+  return(components)
+}

### 2023-05-08 R/minicutter.R
fixing exports

@@ -9,6 +9,7 @@
 #' @import tidyverse
 #' @import Matrix
 #' @import igraph
+#' @export
 
 leafcutter_one_step = function(juncs) {
 

### 2023-05-08 R/plot_isoforms.R

@@ -6,6 +6,7 @@
 #' @examples
 #' plot_isoforms("ENSG00000100320")
 #' @name plot_isoforms
+#' @export
 
 load(file='data/iso_exon_data.rda')
 

### 2023-05-08 R/isoviz_coords.R
add inst folder for files we want to preload

@@ -11,6 +11,9 @@
 #' isoviz_coords()
 #' @name isoviz_coords
 #' @import data.table
+#' @import tidyr
+#' @import dplyr
+#' @import readr
 #' @export
 
 isoviz_coords = function(file_path, gene_trans,

### 2023-05-08 R/isoviz_plot_juncs_to_iso.R
add inst folder for files we want to preload

@@ -1,0 +1,137 @@
+#' A function for obtaining a comprehensive visualization of gene transcript
+#' isoforms with junction coverage across intron clusters information from RNA-seq
+#' and leafcutter clustering
+#' @param gene ensembl ID for now
+#' @param cell_type cell type of interest, default=hESC
+#' @param junction_usage minimum junction usage, default is 5
+#' @param include_common_juncs show junctions common to all isoforms?
+#' @param filter_expressed_isoforms include only isoforms that are detected
+#' @param filter_all_100percent filter junctions that are always included (is this right?)
+#' @return plot
+#' @examples
+#' isoviz_plot_juncs_to_iso()
+#' @name isoviz_plot_juncs_to_iso
+#' @import data.table
+#' @export
+
+isoviz_plot_juncs_to_iso = function(gene = "ENSG00000099875", cell_type = "hESC", junction_usage = 5, include_common_juncs = FALSE, filter_expressed_isoforms = TRUE, filter_all_100percent = FALSE){
+
+  # call map_junction function here
+  df = map_all_junctions(gene, cell_type)
+
+  # filter here based on input parameters- also unknown junctions that belong to common junctions
+  if(include_common_juncs == FALSE){
+    remove_cluster = df %>% filter(junction_category == "common") %>% select(cluster_idx) %>% distinct()
+    df %<>% filter(!cluster_idx %in% remove_cluster$cluster_idx)
+  }
+
+  if(filter_expressed_isoforms == TRUE){
+    df %<>% filter(isoform_expressed == "Likely")
+
+    # create list of expressed transcripts
+    expressed_isoforms = df %>% filter(isoform_expressed == "Likely", transcript_isoforms != "unknown") %>%
+      separate(col = transcript_isoforms, into = c("name", "transcript")) %>% select(transcript) %>% distinct() %>%
+      arrange(transcript) %>%
+      mutate(isoforms = paste0(transcript, collapse = ",")) %>% select(isoforms) %>% distinct()
+  }
+
+  to_remove = df %>% filter(junc.per.cluster == 2 & (junc.usage < junction_usage | junc.usage > (100 - junction_usage)))
+  df %<>% filter(!junc_id %in% to_remove$junc_id, junc.counts > 0)
+
[…]

### 2023-05-08 R/map_all_junctions.R

@@ -1,0 +1,79 @@
+#' A function for mapping junctions to transcript isoforms
+#' @param gene ensembl ID for now
+#' @param cell_type cell type of interest, default=hESC
+#' @return plot
+#' @examples
+#' isoviz_plot_juncs_to_iso()
+#' @name isoviz_plot_juncs_to_iso
+#' @import data.table
+#' @export
+
+# map junctions function
+map_all_junctions = function(gene = "ENSG00000100320", cell_type = "hESC", intron_coords, leafcutter_input){
+
+  # join intron_coords with leafcutter data and mark any unannotated junctions
+
+  # since leafcutter junctions dont have gene annotations, need to combine here first
+  juncs_recluster %<>% select(everything(), junc.counts = readcount) %>%
+    group_by(cluster_idx) %>% mutate(cluster.counts = sum(junc.counts)) %>%
+    ungroup()
+
+  gene_cluster = gene_iso_data %>%
+    left_join(juncs_recluster, by = c("chr" = "chrom", "junc_start" = "start", "junc_end" = "end", "strand")) %>%
+    arrange(cluster_idx) # 24
+
+  # check minicutter output
+  juncs_recluster %>% filter(cluster_idx == "134447")
+
+  name = unique(gene_cluster$gene_name)
+  gene_strand = unique(gene_cluster$strand)
+
+  # to get information on leafcutter junctions that don't match annotation
+  all_gene_clusters = juncs_recluster %>% filter(cluster_idx %in% gene_cluster$cluster_idx) %>%
+    left_join(gene_iso_data, by = c("chrom" = "chr", "start" = "junc_start", "end" = "junc_end", "strand")) %>%
+    filter(is.na(gene_id)) %>%
+    mutate(gene_name = unique(gene_cluster$gene_name), strand = unique(gene_cluster$strand),
+           gene_id = unique(gene_cluster$gene_id), junction_category = "unknown",
+           transcript_isoforms = "unknown", junc_id = paste0("unk.", row_number())) %>%
+    select(chr = chrom, junc_start = start, junc_end = end, everything()) %>%
+    bind_rows(gene_cluster) %>%
+    mutate(junc.counts = ifelse(is.na(cluster_idx), 0, junc.counts), cluster.counts = ifelse(is.na(cluster_idx), 0, cluster.counts)) %>%
+    arrange(cluster_idx)
+
+  n = all_gene_clusters %>% distinct(cluster_idx, junc_id) %>% group_by(cluster_idx) %>%
[…]

### 2023-05-08 R/isoviz_minicutter.R
added minicutter functions

@@ -1,0 +1,132 @@
+#' Helper function to plot cluster sizes
+#' plot_cluster_sizes
+#' This function takes a list of cluster sizes and plots a bar chart of the sizes.
+#'
+#' @param juncs A list of integers representing the sizes of each cluster
+#'
+#' @return A plot
+#'
+#' @examples
+#' plot_cluster_sizes(juncs)
+#'
+
+plot_cluster_sizes = function(juncs) {
+  cluster_sizes = juncs %>% group_by(cluster_idx) %>% summarize(n = n()) %>% ungroup()
+  ta = table(cluster_sizes$n)
+  tibble(
+    cluster_size = as.numeric(names(ta)),
+    num_clusters = as.numeric(ta)) %>%
+    ggplot(aes(cluster_size, num_clusters)) + geom_point() + xlab("cluster size") + ylab("number clusters") + scale_y_log10()
+}
+
+#' Helper function to get junction usage ratios
+#' calculate_usage_ratios
+#' This function takes
+#'
+#' @param juncs A list of integers representing the sizes of each cluster
+#'
+#' @return A plot
+#'
+#' @examples
+#' calculate_usage_ratios(juncs)
+#'
+
+calculate_usage_ratios = function(juncs) {
+    juncs %>% group_by(cluster_idx) %>%
+    mutate(usage_ratio = readcount / sum(readcount)) %>%
+    ungroup() # calculate usage ratios
+}
+
+#' leafcutter_one_step function
+#' A function for obtaining leafcutter intron clusters
+#'
+#' Inputs junction coordinates from regtools
+#' @param juncs_file Gencode gtf or psl file
+#' @return intron clusters
+#' @examples
+#' leafcutter_one_step()
+#' @name leafcutter_one_step
+#' @import tidyverse
+#' @import Matrix
+#' @import igraph
+#' @export
+
+leafcutter_one_step = function(juncs) {
+
+  juncs = juncs %>% select("chrom", "strand", "start", "end", "name", "readcount")
+  splice_sites = bind_rows(
+    juncs %>% select(chrom, strand, position = start),
+    juncs %>% select(chrom, strand, position = end)) %>%
+    distinct() %>%
+    arrange(chrom, strand, position) %>%
+    mutate(idx = 1:n())
+
+  juncs = juncs %>%
+    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", start = "position")) %>%
[…]

### 2023-05-08 R/minicutter.R

@@ -1,41 +1,0 @@
-#' A function for obtaining leafcutter intron clusters
-#'
-#' Inputs junction coordinates from regtools
-#' @param juncs Gencode gtf or psl file
-#' @return intron clusters
-#' @examples
-#' minicutter()
-#' @name minicutter
-#' @import tidyverse
-#' @import Matrix
-#' @import igraph
-#' @export
-
-leafcutter_one_step = function(juncs) {
-
-  splice_sites = bind_rows(
-    juncs %>% select(chrom, strand, position = start),
-    juncs %>% select(chrom, strand, position = end)) %>%
-    distinct() %>%
-    arrange(chrom, strand, position) %>%
-    mutate(idx = 1:n())
-
-  juncs = juncs %>%
-    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", start = "position")) %>%
-    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", end = "position"), suffix = c("_start","_end"))
-
-  nss = nrow(splice_sites)
-
-  intron_connectivity <- sparseMatrix(
-    i = juncs$idx_start,
-    j = juncs$idx_end,
-    dims = c(nss, nss),
-    x = 1,
-    symmetric = TRUE
-  )
-
-  g = graph_from_adjacency_matrix(intron_connectivity, "undirected")
-
-  components(g)$membership[juncs$idx_start]
-  return(components)
-}

### 2023-05-08 temp_code/minicutter.R

@@ -29,7 +29,6 @@ leafcutter_one_step = function(juncs) {
   )
 
   g = graph_from_adjacency_matrix(intron_connectivity, "undirected")
-
   components(g)$membership[juncs$idx_start]
 }
 

### 2023-05-08 R/isoviz_minicutter.R
added minicutter functions

@@ -114,7 +114,8 @@ isoviz_minicutter = function(juncs_file, min_junction_reads=1, plot_summary=TRUE
 
   if(plot_summary){
     print("Printing summary of intron-cluster sizes = how many junctions across in clusters")
-    plot_cluster_sizes(juncs)
+    p = plot_cluster_sizes(juncs)
+    print(p)
   }
 
   juncs = juncs %>% calculate_usage_ratios()
@@ -125,7 +126,8 @@ isoviz_minicutter = function(juncs_file, min_junction_reads=1, plot_summary=TRUE
 
   if(plot_summary){
     print("Printing summary of intron-cluster sizes = how many junctions across in clusters")
-    plot_cluster_sizes(juncs_filtered) + ggtitle("Post cluster refinemenet and removing lowly used junctions")
+    p=plot_cluster_sizes(juncs_filtered) + ggtitle("Post cluster refinemenet and removing lowly used junctions")
+    print(p)
   }
 
   return(juncs)

### 2023-05-08 isoviz_logo.R
added minicutter functions

@@ -1,0 +1,12 @@
+library(hexSticker)
+
+# get image
+#file_path <- system.file("figures", "rbfox2_hesc_filtered.png", package="isoviz")
+file_path <- "/Users/kisaev/Documents/github/isoviz/inst/figures/rbfox2_hesc_filtered.png"
+
+sticker(file_path,
+             package="isoviz",
+             p_color = "orange",
+             p_size = 30,
+             h_fill="white", h_color="#f39c12",
+             filename="inst/figures/logo.png")

### 2023-06-05 R/isoviz_map_junctions.R
add scaling and mapping

@@ -1,65 +1,74 @@
 #' A function for mapping junctions to transcript isoforms
-#' @param gene ensembl ID for now
+#' @param intron_coords description
+#' @param leafcutter_input output from running isoviz_minicutter
 #' @param cell_type cell type of interest, default=hESC
-#' @return plot
+#' @return mapping
 #' @examples
 #' isoviz_plot_juncs_to_iso()
 #' @name isoviz_plot_juncs_to_iso
 #' @import data.table
 #' @export
 
+library(data.table)
+library(plyr)
+library(GenomicFeatures)
+library(tidyr)
+
 # map junctions function
-map_all_junctions = function(gene = "ENSG00000100320", cell_type = "hESC", intron_coords, leafcutter_input){
+map_all_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_input){
 
   # join intron_coords with leafcutter data and mark any unannotated junctions
+  juncs_recluster = leafcutter_input
+
+  # intron_coords rename columns
+  gene_iso_data = gene_intron_coords
+
+  juncs_recluster %<>% dplyr::select(everything(), junc.counts = readcount) %>% group_by(cluster_idx) %>% mutate(cluster.counts = sum(junc.counts)) %>% ungroup()
 
   # since leafcutter junctions dont have gene annotations, need to combine here first
-  juncs_recluster %<>% select(everything(), junc.counts = readcount) %>%
-    group_by(cluster_idx) %>% mutate(cluster.counts = sum(junc.counts)) %>%
-    ungroup()
-
   gene_cluster = gene_iso_data %>%
-    left_join(juncs_recluster, by = c("chr" = "chrom", "junc_start" = "start", "junc_end" = "end", "strand")) %>%
+    left_join(juncs_recluster, by = c("chr" = "chrom", "intron_starts" = "start", "intron_ends" = "end", "strand")) %>%
     arrange(cluster_idx) # 24
 
-  # check minicutter output
-  juncs_recluster %>% filter(cluster_idx == "134447")
-
   name = unique(gene_cluster$gene_name)
+  gene_id = unique(gene_cluster$gene_id)
   gene_strand = unique(gene_cluster$strand)
 
   # to get information on leafcutter junctions that don't match annotation
   all_gene_clusters = juncs_recluster %>% filter(cluster_idx %in% gene_cluster$cluster_idx) %>%
-    left_join(gene_iso_data, by = c("chrom" = "chr", "start" = "junc_start", "end" = "junc_end", "strand")) %>%
[…]

### 2023-06-05 R/isoviz_rescale_introns.R

@@ -1,0 +1,238 @@
+#' A function for reducing the length of introns when visualizing isoforms
+#' through re-scaling all coordinates in the gene.
+#'
+#' Re-scaling intron coordinates while perserving exon alignment by only
+#' impacting the intron width and not the exons. These re-scaled coordinates are
+#' meant only for plotting!
+#'
+#' @param introns #description
+#' @param exons #description
+#' @param width_rescale #description
+#' @param transcripts_group Column name containing transcript name
+#' @return 'data.table()' containing the re-scaled coordinates of 'introns'
+#' and 'exons' for all transcripts in input
+#' @examples
+#' isoviz_rescale_introns()
+#' @name isoviz_rescale_introns
+#' @import data.table
+#' @import tidyr
+#' @import dplyr
+#' @import readr
+#' @export
+
+library(data.table)
+library(plyr)
+library(GenomicFeatures)
+
+#introns = rbfox2_introns
+#exons = rbfox2_exons
+
+isoviz_rescale_introns = function(introns, exons,
+                          width_rescale=50) {
+
+  # very important to consider strand!
+  strand = introns$strand[1]
+  print(strand)
+
+  # order transcripts to figure out which is the reference/starting transcript
+  start_pos = unique(exons[,c("start", "transcript_name")])[order(start)][1]
+  starts = unique(exons[,c("start", "end", "transcript_name", "transcript_length")])[order(start)]
+
+  if(strand == "-"){
+    new_ends = introns$intron_starts
+    new_starts = introns$intron_ends
+    introns$intron_starts = new_starts
+    introns$intron_ends = new_ends
+    new_ends_ex = exons$blockstarts
+    new_starts_ex = exons$blockends
+    exons$blockstarts = new_starts_ex
+    exons$blockends = new_ends_ex
+    new_ends = exons$start
+    new_starts = exons$end
+    exons$start = new_starts
+    exons$end = new_ends
+    starts = unique(exons[,c("start", "end", "transcript_name", "transcript_length")])[order(-start)]
+    start_pos = unique(exons[,c("start", "end", "transcript_name")])[order(-start)][1]
+  }
+
+  introns$length = abs(introns$intron_starts-introns$intron_ends)
+  introns$rescaled_length = introns$length / width_rescale
[…]

### 2023-06-05 examples/testing_plotting_rescaled.R

@@ -1,0 +1,37 @@
+gene_data = rescaled_coords
+
+# setting parameters for the plot based on length of the gene and number of isoforms
+length = max(gene_data$transcript_length) 
+min_start = min(gene_data$new_e_start)
+#min_start=0
+n = gene_data %>% group_by(transcript_name) %>% tally() %>% nrow()
+y_max = n + 1
+max_end = max(gene_data$new_e_end)
+
+# order the isoforms by length for plotting
+order_plot = gene_data %>% group_by(transcript_name, transcript_length) %>% tally() %>%
+  dplyr::arrange(desc(transcript_length), desc(n))
+order_plot$trans_order = 1:nrow(order_plot)
+order_plot = order_plot %>% dplyr::select(transcript_name, trans_order)
+
+to_plot_ordered = gene_data %>% full_join(order_plot, by = c("transcript_name")) %>%
+  arrange(trans_order, new_e_start) %>%
+  mutate(y1 = trans_order - 0.2, y2 = trans_order + 0.2)
+to_plot_ordered = to_plot_ordered %>% mutate(seg_end = new_e_end)
+  
+# make plot
+p1 = ggplot() +
+    geom_segment(aes(x = to_plot_ordered$new_e_start, y = to_plot_ordered$trans_order,
+                     xend = to_plot_ordered$new_e_end, yend = to_plot_ordered$trans_order)) +
+    geom_rect(data = to_plot_ordered, mapping = aes(xmin = new_e_start, xmax = new_e_end, ymin = y1, ymax = y2)) +
+    xlim(min_start, max_end) + ylim(0,y_max) +
+    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$transcript_name)) +
+    theme_bw() + ylab("") +
+    theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "top") +
+    xlab("Re-scaled coordinates") +
+    ggtitle(paste0(" (", to_plot_ordered$strand[1], ")")) +
+    theme(axis.text = element_text(size = 10), plot.title = element_text(hjust = 0.5)) +
+    coord_cartesian(clip = 'off') +
+    theme(plot.margin = margin(0.5,1.5,0.5,0.5, "in"))
+
+print(p1)

### 2023-06-05 R/isoviz_map_junctions.R
updating data

@@ -1,11 +1,12 @@
 #' A function for mapping junctions to transcript isoforms
-#' @param intron_coords description
+#' @param gene_intron_coords obtained intron coordinates for gene of interest
 #' @param leafcutter_input output from running isoviz_minicutter
 #' @param cell_type cell type of interest, default=hESC
+#' @param gencode_intron_all_data expanded annotations of introns/transcripts
 #' @return mapping
 #' @examples
-#' isoviz_plot_juncs_to_iso()
-#' @name isoviz_plot_juncs_to_iso
+#' isoviz_map_junctions()
+#' @name isoviz_map_junctions
 #' @import data.table
 #' @export
 
@@ -15,7 +16,7 @@ library(GenomicFeatures)
 library(tidyr)
 
 # map junctions function
-map_all_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_input){
+isoviz_map_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_input, gencode_intron_all_data){
 
   # join intron_coords with leafcutter data and mark any unannotated junctions
   juncs_recluster = leafcutter_input
@@ -29,6 +30,10 @@ map_all_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_
   gene_cluster = gene_iso_data %>%
     left_join(juncs_recluster, by = c("chr" = "chrom", "intron_starts" = "start", "intron_ends" = "end", "strand")) %>%
     arrange(cluster_idx) # 24
+
+  # get additional columns from the gencode_intron_all_data file
+  gene_cluster = gene_cluster %>%
+    left_join(gencode_intron_all_data, by=c("chr" = "chr", "intron_starts" = "junc_start", "intron_ends"="junc_end", "strand", "gene_id", "gene_name", "gene_type"))
 
   name = unique(gene_cluster$gene_name)
   gene_id = unique(gene_cluster$gene_id)
@@ -59,16 +64,14 @@ map_all_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_
   # of technical issues so only eliminate isoforms where none of their unique
   # junctions have counts, avoid filtering on partially unique for now
 
-  # assign junction categories, fully_unique, partial_unique, common...?
-
   non_expressed_genes = all_gene_clusters %>%
-    separate_longer_delim(transcript_name, delim = ",") %>% group_by(transcript_name, junction_category) %>%
[…]

### 2023-06-05 R/isoviz_plot_juncs_to_iso.R

@@ -1,12 +1,11 @@
 #' A function for obtaining a comprehensive visualization of gene transcript
 #' isoforms with junction coverage across intron clusters information from RNA-seq
 #' and leafcutter clustering
-#' @param gene ensembl ID for now
+#' @param mapped_junctions output from running isoviz_map_junctions
 #' @param cell_type cell type of interest, default=hESC
-#' @param junction_usage minimum junction usage, default is 5
-#' @param include_common_juncs show junctions common to all isoforms?
-#' @param filter_expressed_isoforms include only isoforms that are detected
-#' @param filter_all_100percent filter junctions that are always included (is this right?)
+#' @param junction_usage minimum junction usage as a percentage, default is 5
+#' @param include_all_juncs show all junctions on plot, will still filter out based on junction usage parameter, default is TRUE
+#' @param include_specific_junctions provide list of junctions to plot in second panel, can only specify if include_all_juncs is FALSE
 #' @return plot
 #' @examples
 #' isoviz_plot_juncs_to_iso()
@@ -14,43 +13,40 @@
 #' @import data.table
 #' @export
 
-isoviz_plot_juncs_to_iso = function(gene = "ENSG00000099875", cell_type = "hESC", junction_usage = 5, include_common_juncs = FALSE, filter_expressed_isoforms = TRUE, filter_all_100percent = FALSE){
+library(data.table)
+library(dplyr)
+library(tidyr)
+library(magrittr)
+library(ggsci)
+library(cowplot)
 
-  # call map_junction function here
-  df = map_all_junctions(gene, cell_type)
+isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
+                                    cell_type = "hESC",
+                                    junction_usage = 5,
+                                    include_all_juncs = TRUE,
+                                    include_specific_junctions = c()){
 
-  # filter here based on input parameters- also unknown junctions that belong to common junctions
-  if(include_common_juncs == FALSE){
-    remove_cluster = df %>% filter(junction_category == "common") %>% select(cluster_idx) %>% distinct()
[…]

### 2023-06-05 R/isoviz_plot_juncs_to_iso.R
updating data

@@ -11,6 +11,9 @@
 #' isoviz_plot_juncs_to_iso()
 #' @name isoviz_plot_juncs_to_iso
 #' @import data.table
+#' @import magrittr
+#' @import ggsci
+#' @import cowplot
 #' @export
 
 library(data.table)

### 2023-06-05 R/plot_isoforms.R
updating data

@@ -1,51 +1,0 @@
-#' A function for plotting transcript isoforms for a specific gene
-#'
-#' This function allows you to visualize all annotated isoforms for a given gene.
-#' @param gene Input your gene of interest using ENSG for now. Defaults to "ENSG00000100320" (RBFOX2)
-#' @return A plot.
-#' @examples
-#' plot_isoforms("ENSG00000100320")
-#' @name plot_isoforms
-#' @export
-
-load(file='data/iso_exon_data.rda')
-
-plot_isoforms = function(gene="ENSG00000100320"){
-
-  gene_trans = which(str_detect(genomedata$gene_id, gene))
-  gene_data = genomedata[gene_trans,]
-
-  # setting parameters for the plot based on length of the gene and number of isoforms
-  length = max(gene_data$end) - min(gene_data$start)
-  min_start = min(gene_data$start)
-  n = gene_data %>% group_by(trans_id) %>% tally() %>% nrow()
-  y_max = n + 1
-  max_end = max(gene_data$end)
-
-  # order the isoforms by length for plotting
-  order_plot = gene_data %>% group_by(trans_id, transcript_length) %>% tally() %>%
-    dplyr::arrange(desc(transcript_length), desc(n))
-  order_plot$trans_order = 1:nrow(order_plot)
-  order_plot = order_plot %>% dplyr::select(trans_id, trans_order)
-
-  to_plot_ordered = gene_data %>% full_join(order_plot, by = c("trans_id")) %>%
-    arrange(trans_order, blockstarts) %>%
-    mutate(y1 = trans_order - 0.2, y2 = trans_order + 0.2)
-  to_plot_ordered = to_plot_ordered %>% mutate(seg_end = end)
-
-  # make plot
-  p1 = ggplot() +
-    geom_segment(aes(x = to_plot_ordered$start, y = to_plot_ordered$trans_order,
-                     xend = to_plot_ordered$seg_end, yend = to_plot_ordered$trans_order)) +
-    geom_rect(data = to_plot_ordered, mapping = aes(xmin = blockstarts, xmax = blockends, ymin = y1, ymax = y2)) +
-    xlim(min_start, max_end) + ylim(0,y_max) +
-    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = 0, label = to_plot_ordered$trans_id)) +
-    theme_bw() + ylab("") +
[…]

### 2023-06-25 R/isoviz_map_junctions.R
fixing random bugs

@@ -18,43 +18,49 @@ library(tidyr)
 # map junctions function
 isoviz_map_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_input, gencode_intron_all_data){
 
+  # note make sure to use the junction IDs from the gencode_intron_all_data file
   # join intron_coords with leafcutter data and mark any unannotated junctions
   juncs_recluster = leafcutter_input
+  juncs_recluster$name = NULL
+  print("removing junction IDs obtained from clustering introns, using default names from gencode file")
 
   # intron_coords rename columns
   gene_iso_data = gene_intron_coords
 
-  juncs_recluster %<>% dplyr::select(everything(), junc.counts = readcount) %>% group_by(cluster_idx) %>% mutate(cluster.counts = sum(junc.counts)) %>% ungroup()
+  # get total counts for clusters across their junctions
+  juncs_recluster %<>% dplyr::select(everything(), junc.counts = readcount) %>% dplyr::group_by(cluster_idx) %>% dplyr::mutate(cluster.counts = sum(junc.counts)) %>% ungroup()
 
   # since leafcutter junctions dont have gene annotations, need to combine here first
   gene_cluster = gene_iso_data %>%
     left_join(juncs_recluster, by = c("chr" = "chrom", "intron_starts" = "start", "intron_ends" = "end", "strand")) %>%
-    arrange(cluster_idx) # 24
+    arrange(cluster_idx)
 
   # get additional columns from the gencode_intron_all_data file
   gene_cluster = gene_cluster %>%
     left_join(gencode_intron_all_data, by=c("chr" = "chr", "intron_starts" = "junc_start", "intron_ends"="junc_end", "strand", "gene_id", "gene_name", "gene_type"))
 
+  # the column 'transcript_isoforms' has pre-mapped transcript isoforms that map those junctions/introns
   name = unique(gene_cluster$gene_name)
   gene_id = unique(gene_cluster$gene_id)
   gene_strand = unique(gene_cluster$strand)
 
-  # to get information on leafcutter junctions that don't match annotation
-  all_gene_clusters = juncs_recluster %>% filter(cluster_idx %in% gene_cluster$cluster_idx) %>%
+  # to get information on leafcutter junctions that don't match annotation, still need to plot them
[…]

### 2023-06-25 R/isoviz_plot_juncs_to_iso.R

@@ -24,8 +24,11 @@ library(ggsci)
 library(cowplot)
 
 isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
+                                    gene_introns,
                                     cell_type = "hESC",
-                                    junction_usage = 5,
+                                    junc_usage = 5,
+                                    intron_scale = "no",
+                                    intron_scale_width = 10,
                                     include_all_juncs = TRUE,
                                     include_specific_junctions = c()){
 
@@ -37,12 +40,12 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
     if(length(include_specific_junctions) == 0){
       stop("Must include list of junctions if include_all_juncs == FALSE")
     } else{
-      df %<>% filter(name %in% include_specific_junctions)
+      df %<>% filter(junc_id %in% include_specific_junctions)
     }
   }
 
   # filter based on junction usage
-  df %<>% filter(junc.usage >= junction_usage)
+  df = dplyr::filter(df, junc.usage >= junc_usage)
 
   # setting parameters for the plot based on length of the gene and number of isoforms
   length = max(gene_data$end) - min(gene_data$start)
@@ -76,22 +79,22 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
 
   # code to plot leafcutter cluster info underneath isoform level data
   # aggregate isoform information
-  df$junc_id = df$name
-  df$name = NULL
+  #df$junc_id = df$name sticking to the original junc_id from Megan's gencode file
+  #df$name = NULL
   df = unique(df)
 
   text = df %>% tidyr::separate(col = transcript_name, into = c("name", "transcript")) %>%
     mutate(transcript = ifelse(is.na(transcript), "unknown", transcript)) %>%
     group_by(cluster_idx, junc_id) %>%
     arrange(cluster_idx, junc_id, transcript) %>%
-    mutate(isoforms = toString(unique(transcript))) %>%
+    dplyr::mutate(isoforms = toString(unique(transcript))) %>%
     dplyr::select(1:5, junc_id, junction_category, junc.usage, isoforms, junc.per.cluster) %>% dplyr::distinct() %>% ungroup() %>%
-    mutate(text_plot = paste0(junc.counts, " reads (", junc.usage, "%) : ", isoforms), alpha = junc.usage/100)
[…]

### 2023-06-26 R/isoviz_plot_juncs_to_iso.R
updated intron re-scaling plottig

@@ -23,6 +23,8 @@ library(magrittr)
 library(ggsci)
 library(cowplot)
 
+#gene_data=rbfox2_exons
+
 isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
                                     gene_introns,
                                     cell_type = "hESC",
@@ -53,6 +55,29 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
   max_end = max(gene_data$end)
   n_transcripts = gene_data %>% group_by(trans_id) %>% tally() %>% nrow()
   y_max = n_transcripts + 1
+  strand = unique(gene_data$strand)
+
+  if(intron_scale == "yes"){
+    # also remove NA isoform from mapped junctions if we want to scale
+    df = dplyr::filter(df, !(is.na(transcript_isoforms)))
+    old_exons = gene_data
+    # rescaled gene_data
+    scaled_res = isoviz_rescale_introns(gene_introns, gene_data, intron_scale_width)
+    gene_data = scaled_res[[1]]
+    int_g = scaled_res[[2]]
+    gene_data$blockstarts = gene_data$new_e_start
+    gene_data$blockends = gene_data$new_e_end
+
+    # save old exon starts and ends
+    gene_data$old_e_start = gene_data$start
+    gene_data$old_e_end = gene_data$end
+
+    gene_data$start = gene_data$min_start  # transcript start
+    gene_data$end =  gene_data$max_end  # transcript end
+
+    min_start = min(gene_data$start)
+    max_end = max(gene_data$end)
+  }
 
   # order the isoforms by length for plotting
   order_plot = gene_data %>% group_by(transcript_name, transcript_length) %>% tally() %>%
@@ -63,7 +88,7 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
   to_plot_ordered = gene_data %>% full_join(order_plot, by = c("transcript_name")) %>%
     arrange(trans_order, blockstarts) %>% mutate(seg_end = end)
 
-  # make isoform-level plot- currently there is only an option to plot all isoforms
+  # make isoform - level plot
   p1 = ggplot() +
     geom_segment(aes(x = to_plot_ordered$start, y = to_plot_ordered$trans_order,
                      xend = to_plot_ordered$seg_end, yend = to_plot_ordered$trans_order)) +
@@ -74,13 +99,11 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
[…]

### 2023-06-26 R/isoviz_rescale_introns.R

@@ -28,7 +28,7 @@ library(GenomicFeatures)
 #exons = rbfox2_exons
 
 isoviz_rescale_introns = function(introns, exons,
-                          width_rescale=50) {
+                                  width_rescale=50) {
 
   # very important to consider strand!
   strand = introns$strand[1]
@@ -60,7 +60,7 @@ isoviz_rescale_introns = function(introns, exons,
   transcripts_names = unique(introns$transcript_name)
 
   if(strand == "-"){
-  starts = starts[order(-start)]}
+    starts = starts[order(-start)]}
 
   z = which(starts$start == starts$start[1])
   starts$new_starts = ""
@@ -107,57 +107,57 @@ isoviz_rescale_introns = function(introns, exons,
     # of the most upstream transcript
 
     if(ref=="ref"){ # for reference transcript only
-    exons_remake$new_e_start = starts[starts$transcript_name == trans_name]$start - starts$start[1]
-    exons_remake$new_e_start = abs(as.numeric(exons_remake$new_e_start))
-    exons_remake$new_e_end =  exons_remake$new_e_start + exons_remake$blocksizes[1]
-    print("getting ref coordinates")
-    for (i in 2:nrow(exons_remake)){
-      # update the other exons
-      exons_remake$new_e_start[i] = exons_remake$new_e_end[i-1] + introns_shortened$rescaled_length[i-1]
-      exons_remake$new_e_end[i] = exons_remake$new_e_start[i] + exons_remake$blocksizes[i]}
+      exons_remake$new_e_start = starts[starts$transcript_name == trans_name]$start - starts$start[1]
+      exons_remake$new_e_start = abs(as.numeric(exons_remake$new_e_start))
+      exons_remake$new_e_end =  exons_remake$new_e_start + exons_remake$blocksizes[1]
+      print("getting ref coordinates")
+      for (i in 2:nrow(exons_remake)){
+        # update the other exons
+        exons_remake$new_e_start[i] = exons_remake$new_e_end[i-1] + introns_shortened$rescaled_length[i-1]
+        exons_remake$new_e_end[i] = exons_remake$new_e_start[i] + exons_remake$blocksizes[i]}
 
-    exons_remake$new_transcript_length = max(exons_remake$new_e_end) - min(exons_remake$new_e_start)
-    exons_remake$new_start = min(exons_remake$new_e_start)
-    exons_remake$new_end = max(exons_remake$new_e_end)
[…]

### 2023-06-26 R/isoviz_coords.R
start website

@@ -10,10 +10,6 @@
 #' @examples
 #' isoviz_coords()
 #' @name isoviz_coords
-#' @import data.table
-#' @import tidyr
-#' @import dplyr
-#' @import readr
 #' @export
 
 isoviz_coords = function(file_path, gene_trans,

### 2023-06-26 R/isoviz_map_junctions.R

@@ -7,13 +7,12 @@
 #' @examples
 #' isoviz_map_junctions()
 #' @name isoviz_map_junctions
-#' @import data.table
 #' @export
 
-library(data.table)
-library(plyr)
-library(GenomicFeatures)
-library(tidyr)
+#library(data.table)
+#library(plyr)
+#library(GenomicFeatures)
+#library(tidyr)
 
 # map junctions function
 isoviz_map_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_input, gencode_intron_all_data){

### 2023-06-26 R/isoviz_minicutter.R

@@ -46,9 +46,6 @@ calculate_usage_ratios = function(juncs) {
 #' @examples
 #' leafcutter_one_step()
 #' @name leafcutter_one_step
-#' @import tidyverse
-#' @import Matrix
-#' @import igraph
 #' @export
 
 leafcutter_one_step = function(juncs) {
@@ -90,9 +87,6 @@ leafcutter_one_step = function(juncs) {
 #' @examples
 #' isoviz_minicutter()
 #' @name isoviz_minicutter
-#' @import tidyverse
-#' @import Matrix
-#' @import igraph
 #' @export
 
 isoviz_minicutter = function(juncs_file, min_junction_reads=1, plot_summary=TRUE, min_usage_ratio=0.001) {

### 2023-06-26 R/isoviz_plot_juncs_to_iso.R

@@ -10,18 +10,14 @@
 #' @examples
 #' isoviz_plot_juncs_to_iso()
 #' @name isoviz_plot_juncs_to_iso
-#' @import data.table
-#' @import magrittr
-#' @import ggsci
-#' @import cowplot
 #' @export
 
-library(data.table)
-library(dplyr)
-library(tidyr)
-library(magrittr)
-library(ggsci)
-library(cowplot)
+#library(data.table)
+#library(dplyr)
+#library(tidyr)
+#library(magrittr)
+#library(ggsci)
+#library(cowplot)
 
 #gene_data=rbfox2_exons
 

### 2023-06-26 R/isoviz_rescale_introns.R

@@ -14,15 +14,11 @@
 #' @examples
 #' isoviz_rescale_introns()
 #' @name isoviz_rescale_introns
-#' @import data.table
-#' @import tidyr
-#' @import dplyr
-#' @import readr
 #' @export
 
-library(data.table)
-library(plyr)
-library(GenomicFeatures)
+#library(data.table)
+#library(plyr)
+#library(GenomicFeatures)
 
 #introns = rbfox2_introns
 #exons = rbfox2_exons

### 2023-06-26 R/isoviz_coords.R
added functioning example for isoviz_coords

@@ -8,7 +8,14 @@
 #' @param gene_trans Dataframe with ensembl gene and transcript ids and names
 #' @return list of exon and intron coordinates in that order.
 #' @examples
-#' isoviz_coords()
+#' # Example with psl file and pre-loaded gene to transcript conversions
+#' psl_file <- system.file("data", "gencode.v41.basic.annotation.psl", package="isoviz")
+#' gene_trans <- system.file("data", "gencode_v41_gene_transcript_convert.txt", package="isoviz")
+#'
+#' results <- isoviz_coords(psl_file, gene_trans, input_type = "psl")
+#' print(head(results)[1]) #exons coords
+#' print(head(results)[2]) #introns coords
+#'
 #' @name isoviz_coords
 #' @export
 

### 2023-06-26 isoviz_logo.R
fix readme file

@@ -1,12 +1,48 @@
 library(hexSticker)
 
 # get image
-#file_path <- system.file("figures", "rbfox2_hesc_filtered.png", package="isoviz")
-file_path <- "/Users/kisaev/Documents/github/isoviz/inst/figures/rbfox2_hesc_filtered.png"
+file_path <- "/Users/kisaev/Documents/github/isoviz/inst/figures/Readme_RBFOX2.png"
 
-sticker(file_path,
-             package="isoviz",
-             p_color = "orange",
-             p_size = 30,
-             h_fill="white", h_color="#f39c12",
-             filename="inst/figures/logo.png")
+library(ggplot2)
+library(isoviz)
+
+## basic example code
+file_path <- system.file("data", "gencode.v41.basic.annotation.psl", package="isoviz")
+gene_trans <- system.file("data", "gencode_v41_gene_transcript_convert.txt", package="isoviz")
+
+all_coordinates <- isoviz_coords(file_path, gene_trans, input_type="psl") #use default genome .psl file
+
+exon_coords <- all_coordinates[[1]]
+intron_coords <- all_coordinates[[2]]
+
+rbfox2_exons <- filter(exon_coords, gene_name == "RBFOX2")
+rbfox2_introns <- filter(intron_coords, gene_name == "RBFOX2")
+
+rescaled_coords = isoviz_rescale_introns(rbfox2_introns, rbfox2_exons, width_rescale=10)
+
+# load junctions for cell type of interest or input your own
+junctions <- system.file("data", "hESC-MKNK2-G1_v41_basic.junc", package="isoviz")
+
+# run minicutter to get clusters
+intron_clusts <- isoviz_minicutter(juncs_file = junctions)
+print(head(intron_clusts))
+
+intron_annotations <- system.file("data", "gencode_intron_all_data.rda", package="isoviz")
+load(intron_annotations)
+
+mapped_junctions = isoviz_map_junctions(cell_type = "hESC", rbfox2_introns, intron_clusts, gencode_intron_all_data)
+
+p = isoviz_plot_juncs_to_iso(mapped_junctions, rbfox2_exons, rbfox2_introns,
+                         cell_type = "hESC",
+                         junc_usage = 5, #min junc usage to be included
+                         include_all_juncs = FALSE,
+                         include_specific_junctions = c("junc178149",
+                                                        "junc178148",
+                                                        "junc178145"))
+
+sticker(p,
[…]

### 2023-06-28 R/isoviz_plot_juncs_to_iso.R
update readme

@@ -180,7 +180,8 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
 
   if(intron_scale == "yes"){
     p2 = ggplot() +
-      geom_rect(data = introns, mapping = aes(xmin = intron_starts, xmax = intron_ends, ymin = junc.order - 0.2, ymax = junc.order + 0.2, fill = as.character(cluster_idx))) +
+      geom_rect(data = introns, mapping = aes(xmin = intron_starts, xmax = intron_ends, ymin = junc.order - 0.2, ymax = junc.order + 0.2, fill = as.character(cluster_idx)))+
+      geom_text(data = introns, aes(x = intron_ends, y = junc.order, label = text_plot), hjust = "inward", size = 2) +
       scale_fill_manual(values = mycols) +
       xlim(min_start, max_end) + theme_bw() +
       xlab("Re-scaled Genomic Position (bp)") + ylab("") +

### 2023-06-30 R/isoviz_map_junctions.R
Bug fix for redundant isoform info in plotting function

@@ -14,6 +14,9 @@
 #library(GenomicFeatures)
 #library(tidyr)
 
+#gene_intron_coords = gene_introns
+#leafcutter_input = intron_clusts
+
 # map junctions function
 isoviz_map_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_input, gencode_intron_all_data){
 
@@ -34,11 +37,12 @@ isoviz_map_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutt
     left_join(juncs_recluster, by = c("chr" = "chrom", "intron_starts" = "start", "intron_ends" = "end", "strand")) %>%
     arrange(cluster_idx)
 
-  # get additional columns from the gencode_intron_all_data file
+  # get additional columns from the gencode_intron_all_data file- fixed bug here creating redundant isoform name columns
+  gencode_intron_all_data %<>% select(-transcript_isoforms)
   gene_cluster = gene_cluster %>%
     left_join(gencode_intron_all_data, by=c("chr" = "chr", "intron_starts" = "junc_start", "intron_ends"="junc_end", "strand", "gene_id", "gene_name", "gene_type"))
 
-  # the column 'transcript_isoforms' has pre-mapped transcript isoforms that map those junctions/introns
+  # the column 'transcript_name' has pre-mapped transcript isoforms that map those junctions/introns
   name = unique(gene_cluster$gene_name)
   gene_id = unique(gene_cluster$gene_id)
   gene_strand = unique(gene_cluster$strand)
@@ -70,23 +74,22 @@ isoviz_map_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutt
   # junctions have counts, avoid filtering on partially unique for now
 
   non_expressed_genes = all_gene_clusters %>%
-    separate_longer_delim(transcript_isoforms, delim = ",") %>% dplyr::group_by(transcript_isoforms, junction_category) %>%
+    dplyr::group_by(transcript_name, junction_category) %>%
     mutate(isoform_counts = ifelse(junction_category == "fully_unique", sum(junc.counts), junc.counts)) %>%
-    filter(junction_category == "fully_unique" & isoform_counts == 0) %>% distinct(transcript_isoforms) %>% ungroup()
+    filter(junction_category == "fully_unique" & isoform_counts == 0) %>% distinct(transcript_name) %>% ungroup()
 
   uniquely_targetable = all_gene_clusters %>%
[…]

### 2023-06-30 R/isoviz_plot_juncs_to_iso.R

@@ -5,6 +5,7 @@
 #' @param cell_type cell type of interest, default=hESC
 #' @param junction_usage minimum junction usage as a percentage, default is 5
 #' @param include_all_juncs show all junctions on plot, will still filter out based on junction usage parameter, default is TRUE
+#' @param include_specific_isoforms provide list of isoforms to plot in first panel, default is to plot all
 #' @param include_specific_junctions provide list of junctions to plot in second panel, can only specify if include_all_juncs is FALSE
 #' @return plot
 #' @examples
@@ -19,7 +20,8 @@
 #library(ggsci)
 #library(cowplot)
 
-#gene_data=rbfox2_exons
+#gene_data=gene_exons
+#df = output
 
 isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
                                     gene_introns,
@@ -28,12 +30,18 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
                                     intron_scale = "no",
                                     intron_scale_width = 10,
                                     include_all_juncs = TRUE,
+                                    include_specific_isoforms = c(), 
                                     include_specific_junctions = c()){
 
   # call isoviz_map_junctions first and use output from that
   df = mapped_junctions
 
-  # filter here if the user wants to filter for specific junctions
+  # filter here if the user wants to filter for specific isoforms or junctions
+  if(length(include_specific_isoforms) != 0){
+    df %<>% filter(transcript_name %in% include_specific_isoforms)
+    gene_data %<>% filter(transcript_name %in% include_specific_isoforms)
+  }
+  
   if(include_all_juncs == FALSE){
     if(length(include_specific_junctions) == 0){
       stop("Must include list of junctions if include_all_juncs == FALSE")
@@ -55,7 +63,7 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
 
   if(intron_scale == "yes"){
     # also remove NA isoform from mapped junctions if we want to scale
-    df = dplyr::filter(df, !(is.na(transcript_isoforms)))
+    df = dplyr::filter(df, !(is.na(transcript_name)))
     old_exons = gene_data
     # rescaled gene_data
[…]

### 2023-06-30 R/isoviz_rescale_introns.R

@@ -22,10 +22,15 @@
 
 #introns = rbfox2_introns
 #exons = rbfox2_exons
+#exons = gene_data # has filtered isoforms
+#introns = gene_introns # has all isoforms
 
 isoviz_rescale_introns = function(introns, exons,
                                   width_rescale=50) {
-
+  
+  # filter intron isoform data based on exon data
+  introns %<>% filter(transcript_name %in% exons$transcript_name)
+  
   # very important to consider strand!
   strand = introns$strand[1]
   print(strand)

### 2023-07-04 R/isoviz_get_guide_predictions.R
adding guide predicition function

@@ -1,0 +1,84 @@
+#' A function for obtaining junction-level guide RNA efficiency predictions for a specific gene or junction
+#' Importantly, this function only works for annotated junctions in whatever gencode version you are using
+#' @param gene ensembl ID for now
+#' @param cell_type cell type of interest, default=hESC
+#' @param guides_per_junction the number of highest ranking guides to print out, minimum of 1 and maximum of 8, default is 8
+#' @param include_all_juncs print guides for all junctions of the gene, default is TRUE
+#' @param include_specific_junctions provide list of junctions for guide RNA prediction, can only specify if include_all_juncs is FALSE
+#' @param leafcutter_input output from running isoviz_minicutter
+#' @return table
+#' @examples
+#' isoviz_get_guide_predictions()
+#' @name isoviz_get_guide_predictions
+#' @export
+
+# Updated: Function to get guide RNA predictions for specific junctions
+# merge junction info and predictions
+# generate this in advance and have as a table
+#v41_info = read_tsv("~/knowles_lab/cas13_share/gencode_v41_all_junction_guides.txt", col_names = TRUE) %>%
+#  mutate(guide_sequence = as.character(guide_sequence)) # 2,255,040
+#predictions = read_csv("~/knowles_lab/cas13_share/230428_predictions/combined-corrected-predictions-gencode.csv") %>%
+#  select(guide_id, predicted_lfc) # 2,254,874
+
+#gencode_predictions = v41_info %>% left_join(predictions, by = "guide_id")
+#save(gencode_predictions, file = "~/for_Karin/gencode_predictions.rda")
+
+# #' @import gencode_predictions, juncs_recluster
+#require(rempsyc)
+
+junctions = c("junc188584", "junc188587", "junc188588", "junc188585")
+
+isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "hESC",
+                                 guides_per_junction = 8,
+                                 include_all_juncs = TRUE,
+                                 include_specific_junctions = c(),
+                                 leafcutter_input){
+
+  # 1. load most up-to-date gencode_predictions file (this will be updated)
+  gencode_predictions <- system.file("data", "gencode_predictions.rda.gz", package="isoviz")
+  load(gencode_predictions)
+
+  # ranking of prediction score -> values 0 to 1
[…]

### 2023-07-04 R/isoviz_get_guide_predictions.R
adding guide predicition function

@@ -36,6 +36,7 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
 
   # 1. load most up-to-date gencode_predictions file (this will be updated)
   gencode_predictions <- system.file("data", "gencode_predictions.rda.gz", package="isoviz")
+  gencode_predictions <- gzfile(gencode_predictions, "rb")
   load(gencode_predictions)
 
   # ranking of prediction score -> values 0 to 1

### 2023-07-04 R/isoviz_get_guide_predictions.R
update guide prediction function

@@ -1,5 +1,4 @@
 #' A function for obtaining junction-level guide RNA efficiency predictions for a specific gene or junction
-#' Importantly, this function only works for annotated junctions in whatever gencode version you are using
 #' @param gene ensembl ID for now
 #' @param cell_type cell type of interest, default=hESC
 #' @param guides_per_junction the number of highest ranking guides to print out, minimum of 1 and maximum of 8, default is 8
@@ -12,28 +11,13 @@
 #' @name isoviz_get_guide_predictions
 #' @export
 
-# Updated: Function to get guide RNA predictions for specific junctions
-# merge junction info and predictions
-# generate this in advance and have as a table
-#v41_info = read_tsv("~/knowles_lab/cas13_share/gencode_v41_all_junction_guides.txt", col_names = TRUE) %>%
-#  mutate(guide_sequence = as.character(guide_sequence)) # 2,255,040
-#predictions = read_csv("~/knowles_lab/cas13_share/230428_predictions/combined-corrected-predictions-gencode.csv") %>%
-#  select(guide_id, predicted_lfc) # 2,254,874
-
-#gencode_predictions = v41_info %>% left_join(predictions, by = "guide_id")
-#save(gencode_predictions, file = "~/for_Karin/gencode_predictions.rda")
-
-# #' @import gencode_predictions, juncs_recluster
-#require(rempsyc)
-
-junctions = c("junc188584", "junc188587", "junc188588", "junc188585")
-
 isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "hESC",
                                  guides_per_junction = 8,
                                  include_all_juncs = TRUE,
                                  include_specific_junctions = c(),
                                  leafcutter_input){
 
+  print("Loading gencode sgRNA prediction file")
   # 1. load most up-to-date gencode_predictions file (this will be updated)
   gencode_predictions <- system.file("data", "gencode_predictions.rda.gz", package="isoviz")
   gencode_predictions <- gzfile(gencode_predictions, "rb")
@@ -72,14 +56,25 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
   filtered_guide_info %<>% dplyr::select(Cluster = cluster_idx, JuncID = junc_id, Category = junction_category, Transcripts, Counts = junc.counts, Usage,
[…]

### 2023-07-21 R/isoviz_get_guide_predictions.R
Updated parameters for plot and table functions

@@ -2,8 +2,8 @@
 #' @param gene ensembl ID for now
 #' @param cell_type cell type of interest, default=hESC
 #' @param guides_per_junction the number of highest ranking guides to print out, minimum of 1 and maximum of 8, default is 8
-#' @param include_all_juncs print guides for all junctions of the gene, default is TRUE
-#' @param include_specific_junctions provide list of junctions for guide RNA prediction, can only specify if include_all_juncs is FALSE
+#' @param output_format Either return the table as a dataframe or flextable, default=dataframe
+#' @param include_specific_junctions provide list of junctions for guide RNA prediction
 #' @param leafcutter_input output from running isoviz_minicutter
 #' @return table
 #' @examples
@@ -13,9 +13,9 @@
 
 isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "hESC",
                                  guides_per_junction = 8,
-                                 include_all_juncs = TRUE,
                                  include_specific_junctions = c(),
-                                 leafcutter_input){
+                                 leafcutter_input, 
+                                 output_format = "dataframe"){
 
   print("Loading gencode sgRNA prediction file")
   # 1. load most up-to-date gencode_predictions file (this will be updated)
@@ -29,13 +29,9 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
   gencode_predictions %<>% mutate(score = 1 - (predicted_lfc - min_score) / (max_score - min_score)) %>% arrange(score)
 
   # filter here if the user wants to filter for specific junctions
-  if(include_all_juncs == FALSE){
-    if(length(include_specific_junctions) == 0){
-      stop("Must include list of junctions if include_all_juncs == FALSE")
-    } else{
-      filtered_predictions = gencode_predictions %>% filter(junc_id %in% include_specific_junctions)
-    }
-  } else{
+  if(length(include_specific_junctions) != 0){
+    filtered_predictions = gencode_predictions %>% filter(junc_id %in% include_specific_junctions)
+  }else{
     filtered_predictions = gencode_predictions %>% filter(grepl(gene, gene_id))
   }
 
@@ -56,10 +52,10 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
[…]

### 2023-07-21 R/isoviz_plot_juncs_to_iso.R

@@ -4,9 +4,9 @@
 #' @param mapped_junctions output from running isoviz_map_junctions
 #' @param cell_type cell type of interest, default=hESC
 #' @param junction_usage minimum junction usage as a percentage, default is 5
-#' @param include_all_juncs show all junctions on plot, will still filter out based on junction usage parameter, default is TRUE
 #' @param include_specific_isoforms provide list of isoforms to plot in first panel, default is to plot all
-#' @param include_specific_junctions provide list of junctions to plot in second panel, can only specify if include_all_juncs is FALSE
+#' @param include_specific_junctions provide list of junctions to plot in second panel
+#' @param plot_transcripts_only plot the top plot only showing transcript structure, default is FALSE
 #' @return plot
 #' @examples
 #' isoviz_plot_juncs_to_iso()
@@ -29,9 +29,9 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
                                     junc_usage = 5,
                                     intron_scale = "no",
                                     intron_scale_width = 10,
-                                    include_all_juncs = TRUE,
                                     include_specific_isoforms = c(), 
-                                    include_specific_junctions = c()){
+                                    include_specific_junctions = c(),
+                                    plot_transcripts_only = FALSE){
 
   # call isoviz_map_junctions first and use output from that
   df = mapped_junctions
@@ -42,11 +42,8 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
     gene_data %<>% filter(transcript_name %in% include_specific_isoforms)
   }
   
-  if(include_all_juncs == FALSE){
-    if(length(include_specific_junctions) == 0){
-      stop("Must include list of junctions if include_all_juncs == FALSE")
-    } else{
-      df %<>% filter(junc_id %in% include_specific_junctions)
+  if(length(include_specific_junctions) != 0){
+    df %<>% filter(junc_id %in% include_specific_junctions)
     }
   }
 
@@ -93,6 +90,12 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
     arrange(trans_order, blockstarts) %>% mutate(seg_end = end)
 
   # make isoform - level plot
[…]

### 2023-07-21 R/isoviz_plot_juncs_to_iso.R
Update function parameters

@@ -44,8 +44,8 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
   
   if(length(include_specific_junctions) != 0){
     df %<>% filter(junc_id %in% include_specific_junctions)
-    }
   }
+
 
   # filter based on junction usage
   df = dplyr::filter(df, junc.usage >= junc_usage)

### 2023-07-28 R/isoviz_get_guide_predictions.R
Update TIGER predictions

@@ -24,10 +24,10 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
   load(gencode_predictions)
 
   # ranking of prediction score -> values 0 to 1
-  min_score = min(gencode_predictions$predicted_lfc)
-  max_score = max(gencode_predictions$predicted_lfc)
-  gencode_predictions %<>% mutate(score = 1 - (predicted_lfc - min_score) / (max_score - min_score)) %>% arrange(score)
-
+  # min_score = min(gencode_predictions$predicted_lfc)
+  # max_score = max(gencode_predictions$predicted_lfc)
+  # gencode_predictions %<>% mutate(score = 1 - (predicted_lfc - min_score) / (max_score - min_score)) %>% arrange(score)
+  
   # filter here if the user wants to filter for specific junctions
   if(length(include_specific_junctions) != 0){
     filtered_predictions = gencode_predictions %>% filter(junc_id %in% include_specific_junctions)
@@ -43,14 +43,14 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
     dplyr::mutate(cluster.counts = sum(junc.counts), Usage = round(Usage, 2)) %>% ungroup()
 
   all_guide_info = filtered_predictions %>%
-    left_join(juncs_recluster, by = c("chr" = "chrom", "junc_start" = "start", "junc_end" = "end", "strand")) %>%
-    arrange(junc_id, score)
+    left_join(juncs_recluster, by = c("chr" = "chrom", "start", "end", "strand")) %>%
+    arrange(junc_id, TIGER_score)
 
   # filter based on parameter for number of guides to report and change NA in counts to 0
-  filtered_guide_info = all_guide_info %>% dplyr::group_by(junc_id) %>% slice_max(score, n = guides_per_junction) %>% ungroup()
+  filtered_guide_info = all_guide_info %>% dplyr::group_by(junc_id) %>% slice_max(TIGER_score, n = guides_per_junction) %>% ungroup()
   filtered_guide_info$Transcripts <- gsub(paste0(gene_name, "-"), "", filtered_guide_info$transcript_isoforms)
   filtered_guide_info %<>% dplyr::select(Cluster = cluster_idx, JuncID = junc_id, Category = junction_category, Transcripts, Counts = junc.counts, Usage,
-                                  GuideSeq = guide_sequence, Score = score)
[…]

### 2023-07-28 R/isoviz_get_guide_predictions.R
updated predictions file and function and also readme images

@@ -5,6 +5,7 @@
 #' @param output_format Either return the table as a dataframe or flextable, default=dataframe
 #' @param include_specific_junctions provide list of junctions for guide RNA prediction
 #' @param leafcutter_input output from running isoviz_minicutter
+#' @param output_format set to dataframe to obtain a regular dataframe object, to obtain a nice table, choose "nice_table"
 #' @return table
 #' @examples
 #' isoviz_get_guide_predictions()
@@ -19,9 +20,10 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
 
   print("Loading gencode sgRNA prediction file")
   # 1. load most up-to-date gencode_predictions file (this will be updated)
-  gencode_predictions <- system.file("data", "gencode_predictions.rda.gz", package="isoviz")
+  gencode_predictions <- system.file("data", "230728_gencodev41_final_TIGER_predictions_to_compare.rda.gz", package="isoviz")
   gencode_predictions <- gzfile(gencode_predictions, "rb")
   load(gencode_predictions)
+  gencode_predictions = tiger_pred_out
 
   # ranking of prediction score -> values 0 to 1
   # min_score = min(gencode_predictions$predicted_lfc)
@@ -66,7 +68,8 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
   #           title = paste0(gene_name, ": Top ", guides_per_junction, " gRNAs per Junction"))
   if(output_format == "dataframe"){
     table_return = filtered_guide_info
-  }else{
+  }
+  if(output_format == "nice_table"){
     table_return = nice_table(filtered_guide_info,
                               title = paste0(gene_name, ": Top ", guides_per_junction, " gRNAs per Junction"))
   }

### 2023-07-28 R/isoviz_rescale_introns.R

@@ -189,7 +189,6 @@ isoviz_rescale_introns = function(introns, exons,
     nas = nas[order(-exon_num)]
 
     for(i in 1:nrow(nas)){
-      print(i)
       trans = nas$transcript_name[i]
       exons = filter(ref_attach_to, transcript_name == trans) #all the non NA exons
       exon_n = nas$exon_num[i]

### 2023-07-28 data-raw/genomedata.R

@@ -1,39 +1,0 @@
-## code to prepare `genomedata` dataset goes here
-
-# data-raw/genomedata/R
-# code to prepare `genomedata` dataset goes here
-
-require(readr)
-require(tidyverse)
-require(stringr)
-
-mydataset <- fread("data-raw/gencode.v41.basic.annotation.psl")
-
-# Data cleaning code here...
-
-# loop creates one line for each annotated exon,
-#combining the blocksizes column witht the start position column in the psl file
-
-mydataset$gene_id = sapply(mydataset$V10, function(x){strsplit(x, "_")[[1]][2]})
-mydataset$trans_id = sapply(mydataset$V10, function(x){strsplit(x, "_")[[1]][1]})
-
-# Get coordinates for each exon "block"
-mydataset$blocksizes = sapply(mydataset$V19, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
-mydataset$blockstarts = sapply(mydataset$V21, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
-mydataset$strand = mydataset$V9
-mydataset$chr = mydataset$V14
-mydataset$start = mydataset$V16
-mydataset$end = mydataset$V17
-mydataset$transcript_length = mydataset$end - mydataset$start
-
-# Clean up and seperate blocks into rows
-mydataset = mydataset %>% dplyr::select(chr, start, end, trans_id, gene_id,strand, blocksizes, blockstarts,
-                            transcript_length) %>% separate_rows(blocksizes, blockstarts)
-mydataset$blockstarts=as.numeric(mydataset$blockstarts)
-mydataset$blocksizes=as.numeric(mydataset$blocksizes)
-mydataset$blockends = mydataset$blockstarts + mydataset$blocksizes
-
-iso_exon_data = as.data.table(mydataset)
-
-# Save to package
-usethis::use_data(iso_exon_data, overwrite = TRUE)

### 2023-07-28 isoviz_logo.R

@@ -1,48 +1,0 @@
-library(hexSticker)
-
-# get image
-file_path <- "/Users/kisaev/Documents/github/isoviz/inst/figures/Readme_RBFOX2.png"
-
-library(ggplot2)
-library(isoviz)
-
-## basic example code
-file_path <- system.file("data", "gencode.v41.basic.annotation.psl", package="isoviz")
-gene_trans <- system.file("data", "gencode_v41_gene_transcript_convert.txt", package="isoviz")
-
-all_coordinates <- isoviz_coords(file_path, gene_trans, input_type="psl") #use default genome .psl file
-
-exon_coords <- all_coordinates[[1]]
-intron_coords <- all_coordinates[[2]]
-
-rbfox2_exons <- filter(exon_coords, gene_name == "RBFOX2")
-rbfox2_introns <- filter(intron_coords, gene_name == "RBFOX2")
-
-rescaled_coords = isoviz_rescale_introns(rbfox2_introns, rbfox2_exons, width_rescale=10)
-
-# load junctions for cell type of interest or input your own
-junctions <- system.file("data", "hESC-MKNK2-G1_v41_basic.junc", package="isoviz")
-
-# run minicutter to get clusters
-intron_clusts <- isoviz_minicutter(juncs_file = junctions)
-print(head(intron_clusts))
-
-intron_annotations <- system.file("data", "gencode_intron_all_data.rda", package="isoviz")
-load(intron_annotations)
-
-mapped_junctions = isoviz_map_junctions(cell_type = "hESC", rbfox2_introns, intron_clusts, gencode_intron_all_data)
-
-p = isoviz_plot_juncs_to_iso(mapped_junctions, rbfox2_exons, rbfox2_introns,
-                         cell_type = "hESC",
-                         junc_usage = 5, #min junc usage to be included
-                         include_all_juncs = FALSE,
-                         include_specific_junctions = c("junc178149",
-                                                        "junc178148",
-                                                        "junc178145"))
-
-sticker(p,
-        package="isoviz",
-        p_color = "orange",
-        p_size=30, s_x=1.2, s_y=0.5, s_width=2, s_height=1.5,
-        h_fill="white", h_color="#f39c12",
-        filename="inst/figures/isoviz.png")

### 2023-07-28 temp_code/leafcutter_cluster_regtools.py

@@ -1,538 +1,0 @@
-#!/usr/bin/env python
-
-# https://github.com/griffithlab/regtools
-# /home/yangili1/tools/regtools/build/regtools junctions extract -a 8 -i 50 -I 500000 bamfile.bam -o outfile.junc
-# Using regtools speeds up the junction extraction step by an order of magnitude or more
-
-import sys
-import tempfile
-import os
-import gzip
-import shutil
-
-def main(options,libl):
-
-    if options.cluster is None:
-        pool_junc_reads(libl, options)
-        refine_clusters(options)
-
-    sort_junctions(libl, options)
-    merge_junctions(options)
-    get_numers(options)
-
-def pool_junc_reads(flist, options):
-
-    outPrefix = options.outprefix
-    rundir = options.rundir
-    maxIntronLen = int(options.maxintronlen)
-    nochromcheck = options.nochromcheck
-
-    outFile = "%s/%s_pooled"%(rundir,outPrefix)
-
-    chromLst = ["chr%d"%x for x in range(1,23)]+['chrX','chrY']+["%d"%x for x in range(1,23)]+['X','Y']
-    by_chrom = {}
-
-    # check that all files exist
-    for libl in flist:
-        if not os.path.isfile(libl.strip()):
-            raise ValueError("File %s does not exist."%lib)
-
-    for k,libl in enumerate(flist, 1):
-        lib = libl.strip()
-        if options.verbose:
-            sys.stderr.write("scanning %d/%d: %s ...\n"%(k, len(flist), lib))
-
-        if lib.endswith('.gz'):
-            opener = gzip.open(lib, 'rt')
-        else:
-            opener = open(lib, 'r')
-
-        for ln in opener:
-            lnsplit = ln.split()
-            if len(lnsplit) < 6:
-                sys.stderr.write("Error in %s\n" % lib)
-                continue
-            chrom, A, B, dot, counts, strand, rA, rb, rgb, blockCount, blockSize, blockStarts = lnsplit
-            if int(blockCount) > 2:
-                print(ln)
-                continue
-            # regtools -s 0 (unstranded) now puts "?" in strand field when strand is ambiguous
-            if strand == "?": continue
-            if not nochromcheck and (chrom not in chromLst): continue  # excludes ALT contigs etc
-            Aoff, Boff = blockSize.split(",")
-            A, B = int(A)+int(Aoff), int(B)-int(Boff)+1
-
-            if B-A > int(maxIntronLen): continue
[…]

### 2023-07-28 temp_code/minicutter.R

@@ -1,80 +1,0 @@
-# pulled from David's github on 4/20/23
-# edited regtools input on 4/21/23
-
-require(tidyverse)
-require(Matrix)
-require(igraph)
-
-leafcutter_one_step = function(juncs) {
-
-  splice_sites = bind_rows(
-    juncs %>% select(chrom, strand, position = start),
-    juncs %>% select(chrom, strand, position = end)) %>%
-    distinct() %>%
-    arrange(chrom, strand, position) %>%
-    mutate(idx = 1:n())
-
-  juncs = juncs %>%
-    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", start = "position")) %>%
-    left_join(splice_sites, by = c(chrom = "chrom", strand = "strand", end = "position"), suffix = c("_start","_end"))
-
-  nss = nrow(splice_sites)
-
-  intron_connectivity <- sparseMatrix(
-    i = juncs$idx_start,
-    j = juncs$idx_end,
-    dims = c(nss, nss),
-    x = 1,
-    symmetric = TRUE
-  )
-
-  g = graph_from_adjacency_matrix(intron_connectivity, "undirected")
-  components(g)$membership[juncs$idx_start]
-}
-
-leafcutter_refine = function(juncs, min_usage_ratio = 0.001) {
-  juncs_filtered = juncs %>% filter(usage_ratio >= min_usage_ratio)
-  juncs_filtered$cluster_idx = leafcutter_one_step(juncs_filtered)
-  juncs_filtered
-}
-
-calculate_usage_ratios = function(juncs) {
-  juncs %>% group_by(cluster_idx) %>% mutate(usage_ratio = readcount / sum(readcount)) %>% ungroup() # calculate usage ratios
-}
-
-leafcutter_two_step = function(juncs, min_usage_ratio = 0.001) {
-  juncs$cluster_idx = leafcutter_one_step(juncs) # get inital clustering
-  juncs = juncs %>% calculate_usage_ratios()
-  leafcutter_refine(juncs, min_usage_ratio = min_usage_ratio)
-}
-
-plot_cluster_sizes = function(juncs) {
-  cluster_sizes = juncs %>% group_by(cluster_idx) %>% summarize(n = n()) %>% ungroup()
-  ta = table(cluster_sizes$n)
-  tibble(
-    cluster_size = as.numeric(names(ta)),
-    num_clusters = as.numeric(ta)
-  ) %>%
[…]

### 2023-08-09 R/isoviz_coords.R
Add gtf function and update coords function

@@ -6,6 +6,9 @@
 #' meaningful names (ex RBFOX2)
 #' @param file_path Gencode gtf or psl file
 #' @param gene_trans Dataframe with ensembl gene and transcript ids and names
+#' @param input_type Either 'psl' or 'gtf' for input file type. Default is psl.
+#' @param count_file User can provide a read count file per transcript from a long read experiment
+#' @param min_count Require a specific number of reads per transcript. Only works if providing count file. Default is 5. 
 #' @return list of exon and intron coordinates in that order.
 #' @examples
 #' # Example with psl file and pre-loaded gene to transcript conversions
@@ -19,76 +22,115 @@
 #' @name isoviz_coords
 #' @export
 
+
 isoviz_coords = function(file_path, gene_trans,
-                         input_type="psl"){
-
+                         input_type="psl", count_file = "", min_count = 5){
+  if(input_type=="gtf"){
+    isoviz_gtf_to_psl(file_path, psl_output_file = "converted_gtf.psl", chrom_sizes=NULL)
+    genome_data <- fread(psl_output_file)
+  }
+  
   if(input_type=="psl"){
-
     #print(paste("The input file being used is"), file_path)
     genome_data <- fread(file_path)
-    genome_data$gene_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][2]})
-    genome_data$trans_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][1]})
-
-    # Get coordinates for each "block"
-    genome_data$blocksizes = sapply(genome_data$V19, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
-    genome_data$blockstarts = sapply(genome_data$V21, function(x){paste(strsplit(x, ",")[[1]], collapse=",")})
-    genome_data$strand = genome_data$V9
-    genome_data$chr = genome_data$V14
-    genome_data$start = genome_data$V16
-    genome_data$end = genome_data$V17
-    genome_data$transcript_length = genome_data$end - genome_data$start
-
-    # Clean up and seperate blocks into rows
-    # Need to do different adjustments here to get intron coords that will line up with leafcutter
-    genome_data = genome_data %>% dplyr::select(chr, start, end, trans_id, gene_id,strand, blocksizes, blockstarts,
[…]

### 2023-08-09 R/isoviz_get_guide_predictions.R

@@ -41,7 +41,7 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
 
   # leafcutter junction counts minicutter input load
   juncs_recluster = leafcutter_input
-  juncs_recluster %<>% select(everything(), junc.counts = readcount, Usage = usage_ratio) %>% dplyr::group_by(cluster_idx) %>%
+  juncs_recluster %<>% dplyr::select(everything(), junc.counts = readcount, Usage = usage_ratio) %>% dplyr::group_by(cluster_idx) %>%
     dplyr::mutate(cluster.counts = sum(junc.counts), Usage = round(Usage, 2)) %>% ungroup()
 
   all_guide_info = filtered_predictions %>%

### 2023-08-09 R/isoviz_gtf_to_psl.R

@@ -1,0 +1,145 @@
+#' A function for converting gtf to psl 
+#'
+#' Converts a gtf to psl, depending on the output filename extension;
+#' gtf exons need to be grouped by transcript and sorted by coordinate w/in a transcript
+#' this function is adapted from Flair https://github.com/BrooksLabUCSC/flair
+#'
+#' @param gtf_file_path User provided gtf file 
+#' @param psl_output_file Name for the PSL file to be outputted. Default is "converted_gtf.psl"
+#' @param chrom_sizes Chrom sizes file for psl conversion, recommended
+#' @return psl formatted file
+#' @examples
+#' # Example with gtf file 
+#'
+#' @name isoviz_gtf_to_psl
+#' @export
+
+#  library(GenomicFeatures)
+#  library(progress)
+#  library(data.table)
+#  library(R.utils)      
+#  library(dplyr)
+
+# gtf_file_path = "/Users/kisaev/Downloads/gencode.v41.basic.annotation.gtf.gz"
+
+gtf_to_psl = function(gtf_file_path, psl_output_file="converted_gtf.psl", chrom_sizes=NULL){
+  
+  # Read the GTF file using fread
+  gtf_data <- fread(gtf_file_path, header = FALSE)
+  
+  # Filter out lines that start with '#'
+  gtf_data <- gtf_data[!grepl("^#", V1)]  
+  
+  # Extract relevant columns
+  print("Extracting info from gtf file to compile all exon information...")
+  
+  system.time({
+    attributes <- gtf_data$V9
+    trans_id <- gsub('.*"(ENST\\d+\\.\\d+)".*', '\\1', attributes)
+    this_transcript <- sub(".*\"(ENST\\d+\\.\\d+)\".*", "\\1", trans_id)
+    gene_ids <- gsub('.*"(ENSG\\d+\\.\\d+)".*', '\\1', attributes)
+  })
+  
+  # Create an empty data frame to store exon information
+  exons_df <- data.frame(
+    chrom = gtf_data$V1,
+    ty = gtf_data$V3,
+    start = as.integer(gtf_data$V4) - 1,
+    end = as.integer(gtf_data$V5),
+    strand = gtf_data$V7,
+    trans_id = this_transcript,
+    transcript_id = this_transcript, 
+    gene_id = gene_ids
+  )
+  
+  # Filter for ty == "CDS"
+  z <- which(exons_df$ty == "CDS")
+  cds_df <- exons_df[z,]
+
+  print("Filtered to just include CDS entries")
+  
[…]

### 2023-08-09 R/isoviz_map_junctions.R

@@ -38,7 +38,7 @@ isoviz_map_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutt
     arrange(cluster_idx)
 
   # get additional columns from the gencode_intron_all_data file- fixed bug here creating redundant isoform name columns
-  gencode_intron_all_data %<>% select(-transcript_isoforms)
+  gencode_intron_all_data %<>% dplyr::select(-transcript_isoforms)
   gene_cluster = gene_cluster %>%
     left_join(gencode_intron_all_data, by=c("chr" = "chr", "intron_starts" = "junc_start", "intron_ends"="junc_end", "strand", "gene_id", "gene_name", "gene_type"))
 

### 2023-08-09 R/isoviz_minicutter.R

@@ -50,10 +50,10 @@ calculate_usage_ratios = function(juncs) {
 
 leafcutter_one_step = function(juncs) {
 
-  juncs = juncs %>% select("chrom", "strand", "start", "end", "name", "readcount")
+  juncs = juncs %>% dplyr::select("chrom", "strand", "start", "end", "name", "readcount")
   splice_sites = bind_rows(
-    juncs %>% select(chrom, strand, position = start),
-    juncs %>% select(chrom, strand, position = end)) %>%
+    juncs %>% dplyr::select(chrom, strand, position = start),
+    juncs %>% dplyr::select(chrom, strand, position = end)) %>%
     distinct() %>%
     arrange(chrom, strand, position) %>%
     mutate(idx = 1:n())
@@ -102,7 +102,7 @@ isoviz_minicutter = function(juncs_file, min_junction_reads=1, plot_summary=TRUE
     separate(blockSizes, into = c("five.p", "three.p"), sep = ",") %>%
     mutate(five.p = as.integer(five.p), three.p = as.integer(three.p)) %>%
     mutate(start = start + five.p, end = end - three.p + 1) %>%
-    select(chrom = chrom, start, end, name = junc.name, readcount = score, strand = strand)
+    dplyr::select(chrom = chrom, start, end, name = junc.name, readcount = score, strand = strand)
 
   juncs = leafcutter_one_step(juncs)
 

### 2023-08-09 R/isoviz_plot_juncs_to_iso.R

@@ -96,16 +96,19 @@ isoviz_plot_juncs_to_iso = function(mapped_junctions, gene_data,
     plot_title = paste0(df$gene_name[1], " Junction to Isoform Map (" , to_plot_ordered$strand[1], " strand )")
   }
   
+  segment_data <- to_plot_ordered %>%
+    group_by(transcript_name) %>%
+    slice_head(n = 1)
+  
   p1 = ggplot() +
-    geom_segment(aes(x = to_plot_ordered$start, y = to_plot_ordered$trans_order,
-                     xend = to_plot_ordered$seg_end, yend = to_plot_ordered$trans_order)) +
+    geom_segment(data = segment_data, aes(x = start, y = trans_order, xend = seg_end, yend = trans_order)) +
     geom_rect(data = to_plot_ordered, mapping = aes(xmin = blockstarts, xmax = blockends, ymin = trans_order - 0.3, ymax = trans_order + 0.3)) +
     xlim(min_start, max_end) + ylim(0,y_max) +
-    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, hjust = -0.1, label = to_plot_ordered$transcript_name), size = 3, check_overlap = TRUE) +
+    geom_text(aes(x = Inf, y = to_plot_ordered$trans_order, label = to_plot_ordered$transcript_name), hjust = 0, vjust = 0.5, size = 3, check_overlap = TRUE) +
     theme_bw() + ylab("") + xlab("") +
     theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
           axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.position = "top",
-          plot.title = element_text(hjust = 0.5), plot.margin = margin(0.1,1,0,0.1, "in")) +
+          plot.title = element_text(hjust = 0.5), plot.margin = margin(0.1,2,0,0.1, "in")) +
     ggtitle(plot_title) +
     coord_cartesian(xlim = c(min_start, max_end), clip = 'off')
 

### 2023-08-10 R/isoviz_junction_to_transcript_summary.R
added function to get junction uniqueness

@@ -1,0 +1,119 @@
+#' A function for summarizing junction and transcript level information from GENCODE annotated transcripts or long read sequencing transcripts.
+#' This is especially useful to identify unique junctions that are specific to a single transcript.
+#'
+#' Inputs gencode intron rda file or long read intron rda file (output from isoviz_coords), outputs summary table
+#' and summary plot
+#' 
+#' @param iso_intron_data This is output from isoviz_coords.
+#' @param type Either 'gencode' or 'longread' depending on the initial source of transcript information. Default is gencode.
+#' @param plot_path This is optional, but if you would like a summary plot, give the path and name of the output. Include proper extension such as '.pdf' or .'png'. Default is no plot.
+#' @return table summarizing junction and transcript level information.
+#' 
+#' @name isoviz_junction_to_transcript_summary
+#' @export
+
+isoviz_junction_to_transcript_summary = function(iso_intron_data, type = "gencode", plot_path = ""){
+  
+  # check input file
+  if(ncol(iso_intron_data) != 10){
+    stop("Input file has a different number of columns than expected. Columns should be: chr, intron_starts, intron_ends, gene_id, trans_id, strand, gene_name, transcript_name, gene_type, transcript_type")
+  }
+  
+  # filter for protein coding and lncRNA genes
+  iso_intron_pc = iso_intron_data %>% filter(gene_type == "protein_coding" | gene_type == "lncRNA")
+  
+  # filter for protein-coding and lncRNA transcripts- we can only do this for gencode annotations since we wont have this info for novel transcripts in long read data
+  if(type == "gencode"){
+    iso_intron_pc %<>% filter(transcript_type == "protein_coding" | transcript_type == "lncRNA")
+  }
+  
+  # remove transcript ids and collapse on junction coords- this gives the total number of junctions in your annotations
+  # assign universal ids for junction
+  collapsed_junctions = iso_intron_pc %>% dplyr::select(-trans_id, -transcript_name, -gene_type, -transcript_type) %>% 
+    distinct() %>% dplyr::select(chr, start = intron_starts, end = intron_ends, gene_id, gene_name, strand) %>% 
[…]

### 2023-08-11 R/isoviz_coords.R
added function to obtain gene trans conversion from gtf

@@ -81,7 +81,7 @@ isoviz_coords = function(file_path, gene_trans,
   print(paste(length(unique(intron_data$trans_id)), "transcripts with at least one intron!"))
 
   # Convert to gene and transcript names (should this be placed somewhere else?)
-  convert = read_tsv(gene_trans, col_names = TRUE)
+  convert = fread(gene_trans)
 
   # Should avoid selecting columns this way
   #trans_info = genome_data %>% dplyr::select(1, 4, 5, 6) %>% distinct() %>% filter(chr != "chrY", chr != "chrM")
@@ -93,7 +93,7 @@ isoviz_coords = function(file_path, gene_trans,
     left_join(convert, by = c("gene_id", "trans_id", "gene_name", "gene_type")) %>% 
     dplyr::select(chr, intron_starts, intron_ends, gene_id, trans_id, strand, gene_name, transcript_name, gene_type, transcript_type)
     
-  id = intron_data %>% dplyr::select(gene_id, trans_id) %>% distinct() %>% arrange(gene_id, desc(trans_id)) %>%
+  id = intron_data %>% dplyr::select(gene_id, trans_id) %>% distinct() %>% dplyr::arrange(gene_id, desc(trans_id)) %>%
     group_by(gene_id) %>% mutate(id = 1:n()) %>% ungroup()
     
   intron_data$transcript_name[intron_data$transcript_name == ""] <- NA

### 2023-08-11 R/isoviz_get_gene_transcript_conversion.R

@@ -1,0 +1,57 @@
+#' A function for converting gtf to psl 
+#'
+#' Converts a gtf to psl, depending on the output filename extension;
+#' gtf exons need to be grouped by transcript and sorted by coordinate w/in a transcript
+#' this function is adapted from Flair https://github.com/BrooksLabUCSC/flair
+#'
+#' @param gtf_file_path User provided gtf file 
+#' @param output_file Name for the PSL file to be outputted. Default is "converted_gtf.psl"
+#' @return table with gene ids, trancsript ids, names... 
+#' @examples
+#' # Example with gtf file 
+#'
+#' @name isoviz_get_gene_transcript_conversion
+#' @export
+
+isoviz_get_gene_transcript_conversion = function(gtf_file, output_file="gene_transcript_convert.txt"){
+  
+  gtf = read_tsv(gtf_file, col_names = FALSE, skip = 5) %>% dplyr::filter(X3 == "transcript") 
+
+  attributes <- gtf$X9
+  print(length(attributes))
+  
+  # Function to extract specific attributes
+  extract_attributes <- function(attr_string) {
+    attr_list <- strsplit(attr_string, "; ")[[1]]
+    attr_pairs <- sapply(attr_list, function(attr) {
+      key_value <- strsplit(attr, " ", fixed = TRUE)
+      list(key = key_value[[1]][1], value = key_value[[1]][2])
+    })
+    
+    attr_df <- as.data.frame(matrix(unlist(attr_pairs), ncol = 2, byrow = TRUE))
+    colnames(attr_df) <- c("Attribute", "Value")
+    specific_attributes <- c("gene_id", "gene_name", "transcript_id", "transcript_name", "gene_type", "transcript_type")
+    specific_data <- attr_df[attr_df$Attribute %in% specific_attributes, ]
+    specific_data <- spread(specific_data, Attribute, Value)
+    #specific_data <- specific_data %>%
+    #  mutate_all(~ str_replace(., "^\"|\"$", "")) %>% 
+    #  mutate_all(~ str_replace(., "^\"|\"$", ""))
+    # clean up values 
+    return(specific_data)
+  }
+  
+  # Apply the function to all attributes
+  all_extracted <- llply(attributes, extract_attributes, .progress="text")
+  
+  # Combine all extracted attributes into a single data frame
+  final_df <- as.data.table(ldply(all_extracted))
+  final_df <- final_df %>%
[…]

### 2023-08-11 R/isoviz_gtf_to_psl.R

@@ -22,7 +22,7 @@
 
 # gtf_file_path = "/Users/kisaev/Downloads/gencode.v41.basic.annotation.gtf.gz"
 
-gtf_to_psl = function(gtf_file_path, psl_output_file="converted_gtf.psl", chrom_sizes=NULL){
+isoviz_gtf_to_psl = function(gtf_file_path, psl_output_file="converted_gtf.psl", specie="human", chrom_sizes=NULL){
   
   # Read the GTF file using fread
   gtf_data <- fread(gtf_file_path, header = FALSE)
@@ -33,12 +33,17 @@ gtf_to_psl = function(gtf_file_path, psl_output_file="converted_gtf.psl", chrom_
   # Extract relevant columns
   print("Extracting info from gtf file to compile all exon information...")
   
-  system.time({
+  if (specie=="human") {
     attributes <- gtf_data$V9
     trans_id <- gsub('.*"(ENST\\d+\\.\\d+)".*', '\\1', attributes)
     this_transcript <- sub(".*\"(ENST\\d+\\.\\d+)\".*", "\\1", trans_id)
-    gene_ids <- gsub('.*"(ENSG\\d+\\.\\d+)".*', '\\1', attributes)
-  })
+    gene_ids <- gsub('.*"(ENSG\\d+\\.\\d+)".*', '\\1', attributes)}
+  
+  if (specie=="mouse") {
+    attributes <- gtf_data$V9
+    trans_id <- gsub('.*"(ENSMUST\\d+\\.\\d+)".*', '\\1', attributes)
+    this_transcript <- sub(".*\"(ENSMUST\\d+\\.\\d+)\".*", "\\1", trans_id)
+    gene_ids <- gsub('.*"(ENSMUSG\\d+\\.\\d+)".*', '\\1', attributes)}
   
   # Create an empty data frame to store exon information
   exons_df <- data.frame(
@@ -124,8 +129,13 @@ gtf_to_psl = function(gtf_file_path, psl_output_file="converted_gtf.psl", chrom_
     }
     
     qstarts_str <- paste(qstarts, collapse = ",")
+    qstarts_str <- paste(qstarts_str, ",", sep = "")
+    
     blocksizes_str <- paste(blocksizes, collapse = ",")
+    blocksizes_str <- paste(blocksizes_str, ",", sep = "")
+    
     blockstarts_str <- paste(blockstarts, collapse = ",")
+    blockstarts_str <- paste(blockstarts_str, ",", sep = "")
     
[…]

### 2023-08-11 R/isoviz_junction_to_transcript_summary.R

@@ -110,7 +110,7 @@ isoviz_junction_to_transcript_summary = function(iso_intron_data, type = "gencod
             legend.position = "top") +
       guides(fill = guide_legend(title = "", nrow = 2, byrow = TRUE))
     
-    ggsave(plot_path, width = 5, height = 4)
+    ggsave(plot_path, width = 6, height = 4)
     
   }
   

### 2023-09-06 R/isoviz_coords.R
use import.gff to process gtf files

@@ -4,7 +4,7 @@
 #' Inputs gencode gtf or psl file, outputs exon coords (for plotting) and
 #' intron coords (for leafcutter integration) with ENSG/ENST names plus
 #' meaningful names (ex RBFOX2)
-#' @param file_path Gencode gtf or psl file
+#' @param file_path Gencode PSL file! (can be custom PSL file though)
 #' @param gene_trans Dataframe with ensembl gene and transcript ids and names
 #' @param input_type Either 'psl' or 'gtf' for input file type. Default is psl.
 #' @param count_file User can provide a read count file per transcript from a long read experiment
@@ -24,16 +24,10 @@
 
 
 isoviz_coords = function(file_path, gene_trans,
-                         input_type="psl", count_file = "", min_count = 5){
-  if(input_type=="gtf"){
-    isoviz_gtf_to_psl(file_path, psl_output_file = "converted_gtf.psl", chrom_sizes=NULL)
-    genome_data <- fread(psl_output_file)
-  }
+                         count_file = "", min_count = 5){
   
-  if(input_type=="psl"){
-    #print(paste("The input file being used is"), file_path)
-    genome_data <- fread(file_path)
-  }
+  print("Please ensure your input file is a PSL file~")
+  genome_data <- fread(file_path)
   
   genome_data$gene_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][2]})
   genome_data$trans_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][1]})
@@ -54,10 +48,10 @@ isoviz_coords = function(file_path, gene_trans,
   genome_data$blockstarts=as.numeric(genome_data$blockstarts)
   genome_data$blocksizes=as.numeric(genome_data$blocksizes)
   genome_data$blockends = genome_data$blockstarts + genome_data$blocksizes
+  genome_data = as.data.table(genome_data)
+  genome_data = unique(genome_data)
   
   transcript_ids = unique(genome_data$trans_id)
-  print(length(transcript_ids)) # 116,566
-
   print(paste(length(unique(genome_data$trans_id)), "transcripts with at least one intron!"))
   
   # Split the data frame by trans_id
@@ -82,24 +76,30 @@ isoviz_coords = function(file_path, gene_trans,
[…]

### 2023-09-06 R/isoviz_get_gene_transcript_conversion.R

@@ -5,7 +5,7 @@
 #' this function is adapted from Flair https://github.com/BrooksLabUCSC/flair
 #'
 #' @param gtf_file_path User provided gtf file 
-#' @param output_file Name for the PSL file to be outputted. Default is "converted_gtf.psl"
+#' @param output_file Name for the txt file to be outputted. Default is "gene_transcript_convert.txt"
 #' @return table with gene ids, trancsript ids, names... 
 #' @examples
 #' # Example with gtf file 
@@ -15,43 +15,23 @@
 
 isoviz_get_gene_transcript_conversion = function(gtf_file, output_file="gene_transcript_convert.txt"){
   
-  gtf = read_tsv(gtf_file, col_names = FALSE, skip = 5) %>% dplyr::filter(X3 == "transcript") 
+  gr <- import.gff(gtf_file)
+  gtf_data <- as.data.table(gr)
+  print("Keeping only protein-coding and lncRNA genes")
+  gtf_data = dplyr::filter(gtf_data, gene_type %in% c("protein_coding", "lncRNA"))
+  
+  # keep only the following columns 
+  exons_df = gtf_data %>% dplyr::select(gene_id, transcript_id, gene_name,
+                                        transcript_name, gene_type, transcript_type)
+  exons_df = unique(exons_df)
+  
+  # filter rows where transcript_id is NA 
+  exons_df = dplyr::filter(exons_df, !(is.na(transcript_id)))
+  colnames(exons_df)[which(colnames(exons_df) == "transcript_id")] = "trans_id"
 
-  attributes <- gtf$X9
-  print(length(attributes))
-  
-  # Function to extract specific attributes
-  extract_attributes <- function(attr_string) {
-    attr_list <- strsplit(attr_string, "; ")[[1]]
-    attr_pairs <- sapply(attr_list, function(attr) {
-      key_value <- strsplit(attr, " ", fixed = TRUE)
-      list(key = key_value[[1]][1], value = key_value[[1]][2])
-    })
-    
-    attr_df <- as.data.frame(matrix(unlist(attr_pairs), ncol = 2, byrow = TRUE))
-    colnames(attr_df) <- c("Attribute", "Value")
-    specific_attributes <- c("gene_id", "gene_name", "transcript_id", "transcript_name", "gene_type", "transcript_type")
-    specific_data <- attr_df[attr_df$Attribute %in% specific_attributes, ]
-    specific_data <- spread(specific_data, Attribute, Value)
[…]

### 2023-09-06 R/isoviz_gtf_to_psl.R

@@ -7,6 +7,7 @@
 #' @param gtf_file_path User provided gtf file 
 #' @param psl_output_file Name for the PSL file to be outputted. Default is "converted_gtf.psl"
 #' @param chrom_sizes Chrom sizes file for psl conversion, recommended
+#' @param filter_CDS if "yes" then will only include CDS exons in the PSL output file
 #' @return psl formatted file
 #' @examples
 #' # Example with gtf file 
@@ -14,60 +15,50 @@
 #' @name isoviz_gtf_to_psl
 #' @export
 
-#  library(GenomicFeatures)
-#  library(progress)
-#  library(data.table)
-#  library(R.utils)      
-#  library(dplyr)
-
-# gtf_file_path = "/Users/kisaev/Downloads/gencode.v41.basic.annotation.gtf.gz"
-
-isoviz_gtf_to_psl = function(gtf_file_path, psl_output_file="converted_gtf.psl", specie="human", chrom_sizes=NULL){
+isoviz_gtf_to_psl = function(gtf_file_path, 
+                             psl_output_file="converted_gtf.psl", 
+                             specie="human", 
+                             filter_CDS="no",
+                             chrom_sizes=NULL){
   
-  # Read the GTF file using fread
-  gtf_data <- fread(gtf_file_path, header = FALSE)
-  
-  # Filter out lines that start with '#'
-  gtf_data <- gtf_data[!grepl("^#", V1)]  
+  # Read the GTF file 
+  # Use import.gff to read the GFF file into a GRanges object
+  gr <- import.gff(gtf_file_path)
+  gtf_data <- as.data.table(gr)
   
   # Extract relevant columns
-  print("Extracting info from gtf file to compile all exon information...")
+  print("Keeping only protein-coding and lncRNA genes")
+  gtf_data = dplyr::filter(gtf_data, gene_type %in% c("protein_coding", "lncRNA"))
   
-  if (specie=="human") {
-    attributes <- gtf_data$V9
-    trans_id <- gsub('.*"(ENST\\d+\\.\\d+)".*', '\\1', attributes)
-    this_transcript <- sub(".*\"(ENST\\d+\\.\\d+)\".*", "\\1", trans_id)
-    gene_ids <- gsub('.*"(ENSG\\d+\\.\\d+)".*', '\\1', attributes)}
+  # keep only the following columns 
[…]

### 2023-09-06 R/isoviz_gtf_to_psl.R
updated junction files for A375 and HEK293

@@ -17,7 +17,6 @@
 
 isoviz_gtf_to_psl = function(gtf_file_path, 
                              psl_output_file="converted_gtf.psl", 
-                             specie="human", 
                              filter_CDS="no",
                              chrom_sizes=NULL){
   

### 2023-09-06 R/isoviz_map_junctions.R

@@ -1,21 +1,13 @@
 #' A function for mapping junctions to transcript isoforms
 #' @param gene_intron_coords obtained intron coordinates for gene of interest
 #' @param leafcutter_input output from running isoviz_minicutter
-#' @param cell_type cell type of interest, default=hESC
+#' @param cell_type cell type of interest, default=hESC other options are A375 and HEK293
 #' @param gencode_intron_all_data expanded annotations of introns/transcripts
 #' @return mapping
 #' @examples
 #' isoviz_map_junctions()
 #' @name isoviz_map_junctions
 #' @export
-
-#library(data.table)
-#library(plyr)
-#library(GenomicFeatures)
-#library(tidyr)
-
-#gene_intron_coords = gene_introns
-#leafcutter_input = intron_clusts
 
 # map junctions function
 isoviz_map_junctions = function(cell_type = "hESC", gene_intron_coords, leafcutter_input, gencode_intron_all_data){

### 2023-09-06 R/isoviz_coords.R
updated functions to work with various long read formats

@@ -29,6 +29,9 @@ isoviz_coords = function(file_path, gene_trans,
   print("Please ensure your input file is a PSL file~")
   genome_data <- fread(file_path)
   
+  # Set options to avoid scientific notation
+  options(scipen = 999)
+  
   genome_data$gene_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][2]})
   genome_data$trans_id = sapply(genome_data$V10, function(x){strsplit(x, "_")[[1]][1]})
   
@@ -45,6 +48,7 @@ isoviz_coords = function(file_path, gene_trans,
   # Need to do different adjustments here to get intron coords that will line up with leafcutter
   genome_data = genome_data %>% dplyr::select(chr, start, end, trans_id, gene_id,strand, blocksizes, blockstarts,
                                               transcript_length) %>% separate_rows(blocksizes, blockstarts)
+  
   genome_data$blockstarts=as.numeric(genome_data$blockstarts)
   genome_data$blocksizes=as.numeric(genome_data$blocksizes)
   genome_data$blockends = genome_data$blockstarts + genome_data$blocksizes

### 2023-09-06 R/isoviz_get_gene_transcript_conversion.R

@@ -6,6 +6,7 @@
 #'
 #' @param gtf_file_path User provided gtf file 
 #' @param output_file Name for the txt file to be outputted. Default is "gene_transcript_convert.txt"
+#' @param gtf_source "gencode" is default otherwise enter "nongencode" 
 #' @return table with gene ids, trancsript ids, names... 
 #' @examples
 #' # Example with gtf file 
@@ -13,12 +14,23 @@
 #' @name isoviz_get_gene_transcript_conversion
 #' @export
 
-isoviz_get_gene_transcript_conversion = function(gtf_file, output_file="gene_transcript_convert.txt"){
+isoviz_get_gene_transcript_conversion = function(gtf_file, 
+                                                 gtf_source = "gencode",
+                                                 output_file="gene_transcript_convert.txt"){
   
   gr <- import.gff(gtf_file)
   gtf_data <- as.data.table(gr)
-  print("Keeping only protein-coding and lncRNA genes")
-  gtf_data = dplyr::filter(gtf_data, gene_type %in% c("protein_coding", "lncRNA"))
+  
+  if(gtf_source == "gencode"){
+    print("Keeping only protein-coding and lncRNA genes")
+    gtf_data = dplyr::filter(gtf_data, gene_type %in% c("protein_coding", "lncRNA"))}
+  
+  if(gtf_source == "nongencode"){
+    gtf_data$gene_name = gtf_data$gene_id
+    gtf_data$transcript_name = gtf_data$transcript_id
+    gtf_data$gene_type = "protein_coding"
+    gtf_data$transcript_type = "protein_coding"
+  }
   
   # keep only the following columns 
   exons_df = gtf_data %>% dplyr::select(gene_id, transcript_id, gene_name,

### 2023-09-06 R/isoviz_gtf_to_psl.R

@@ -8,6 +8,7 @@
 #' @param psl_output_file Name for the PSL file to be outputted. Default is "converted_gtf.psl"
 #' @param chrom_sizes Chrom sizes file for psl conversion, recommended
 #' @param filter_CDS if "yes" then will only include CDS exons in the PSL output file
+#' @param gtf_source "gencode" is default otherwise enter "nongencode" 
 #' @return psl formatted file
 #' @examples
 #' # Example with gtf file 
@@ -17,7 +18,8 @@
 
 isoviz_gtf_to_psl = function(gtf_file_path, 
                              psl_output_file="converted_gtf.psl", 
-                             filter_CDS="no",
+                             gtf_source = "gencode",
+                             filter_CDS = "no",
                              chrom_sizes=NULL){
   
   # Read the GTF file 
@@ -25,9 +27,14 @@ isoviz_gtf_to_psl = function(gtf_file_path,
   gr <- import.gff(gtf_file_path)
   gtf_data <- as.data.table(gr)
   
+  # Set options to avoid scientific notation
+  options(scipen = 999)
+  
   # Extract relevant columns
   print("Keeping only protein-coding and lncRNA genes")
-  gtf_data = dplyr::filter(gtf_data, gene_type %in% c("protein_coding", "lncRNA"))
+  
+  if(gtf_source == "gencode"){
+    gtf_data = dplyr::filter(gtf_data, gene_type %in% c("protein_coding", "lncRNA"))}
   
   # keep only the following columns 
   exons_df = gtf_data %>% dplyr::select(seqnames, type, start, end, 

### 2023-09-09 R/isoviz_get_guide_predictions.R
updated TIGER predictions file

@@ -20,7 +20,7 @@ isoviz_get_guide_predictions = function(gene = "ENSG00000100320", cell_type = "h
 
   print("Loading gencode sgRNA prediction file")
   # 1. load most up-to-date gencode_predictions file (this will be updated)
-  gencode_predictions <- system.file("data", "230728_gencodev41_final_TIGER_predictions_to_compare.rda.gz", package="isoviz")
+  gencode_predictions <- system.file("data", "230828_gencodev41_final_TIGER_predictions_to_compare.rda.gz", package="isoviz")
   gencode_predictions <- gzfile(gencode_predictions, "rb")
   load(gencode_predictions)
   gencode_predictions = tiger_pred_out
